---
title: "Score screen hits using empirical FDR"
author: "David Wu"
output:
  html_document:
    df_print: paged
---

## Purpose
Determine hits based on 0.05 FDR threshold

## Setup 
Set working directory to project directory
```{r setup}
require(knitr)
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

### Load themes and functions
```{r, message=FALSE}
source('analysis/universal/functions.R') # functions for all project components

source('analysis/universal/themes.R') # themes for all project components

theme_set(theme_publication())

source('analysis/screen/scripts/functions.R')  # functions for this component; may be empty
```

### Directories
```{r}
analysis_dir <- 'analysis/screen/output/02_fdr'
dir.create(analysis_dir, showWarnings = FALSE)
```

### Results directories
```{r}
result_dir <- 'analysis/screen/output/01_process_experiments/results'
result_subdirs <- list.files(result_dir, full.names = TRUE)
names(result_subdirs) <- basename(result_subdirs)
result_subdirs
```

Read in results
```{r, message = FALSE}
results <- map(result_subdirs, function(i) {
  experiment_dirs <- list.files(i, full.names = TRUE)
  names(experiment_dirs) <- basename(experiment_dirs)
  
  analysis <- map(experiment_dirs, function(j) {
    print(c(basename(i), basename(j)))
    screen_analysis(j)
    
  })
  
  replicates <- map(experiment_dirs, function(j) {
    print(c(basename(i), basename(j)))
    screen_replicates(j)
  })
  
  list('analysis' = analysis,
       'replicates' = replicates)
})
```


```{r}
results$Differentiation$analysis$Coding$table
```
## Analysis
### Visualization
```{r}
volcano <- map(results, function(i) {
  map(i$analysis, function(j) {
    j$table %>% screen_volcano()
  })
})
```

```{r, fig.width= 8, fig.height=8}
volcano$Differentiation %>% wrap_plots(ncol = 2)
```

```{r, fig.width= 8, fig.height=8}
volcano$Proliferation %>% wrap_plots(ncol = 2)
```
### Summarize results
```{r}
result_summary <- map(results, function(i) {
  map(i$analysis, function(j) {
    screen_table <- j$table
    assay <- screen_table %>% pull(assay) %>% unique()
    library <- screen_table %>% pull(library) %>% unique()
    experiment <- screen_table %>% pull(experiment) %>% unique()
    
    hits <- screen_table %>% filter(legend == 'Hit Gene') %>% pull(feature_id) %>% n_distinct()
    threshold <- screen_table %>% pull(threshold) %>% unique()
    
    list(assay = assay,
         library = library,
         experiment = experiment,
         hits = hits,
         threshold = threshold)
  }) %>% bind_rows()
}) %>% bind_rows()

result_summary %>% filter(assay == 'Differentiation')
```
```{r}
result_summary %>% filter(assay == 'Proliferation')
```

## Merge
### Merge output into one table
```{r}
differentiaton <- map(results$Differentiation$analysis, function(i) {
  i$table
}) %>% bind_rows()

proliferation <- map(results$Proliferation$analysis, function(i) {
  i$table
}) %>% bind_rows()

full_results <- bind_rows(differentiaton, proliferation) %>% filter(experiment != 'Validation')
full_results
```
### Merge lncRNA data and recalculate screen scores
```{r}
lncrna_results <- full_results %>% filter(library == 'lncRNA')

lncrna_recalculate <- map(c('Differentiation', 'Proliferation'), function(i) {
  lncrna_results %>% 
    filter(assay == i) %>% 
    select(-score, -threshold, -control_sd) %>% 
  screen_calculate_score() %>% 
  screen_mark_hits()
}) %>% bind_rows()

lncrna_recalculate %>% select(assay, threshold) %>% unique()
```

Tally final total hit count
```{r}
lncrna_recalculate %>% group_by(assay, status) %>% tally()
```

Combined volcano plot 
```{r}
map(c('Differentiation', 'Proliferation'), function(i) {
    lncrna_recalculate %>% 
    filter(assay == i) %>% 
    screen_volcano()
}) %>% wrap_plots()
  
```

### Recombine full results
```{r}
full_results <- full_results %>% 
  filter(library == 'Coding') %>% 
  bind_rows(lncrna_recalculate)
full_results
```


### Total hits
```{r}
full_results %>% filter(group == 'Gene', status == 'Hit') %>% pull(feature_id) %>% n_distinct()
```
Coding hits
```{r}
full_results %>% filter(group == 'Gene', status == 'Hit', library == 'Coding') %>% pull(feature_id) %>% n_distinct()
```

lncRNA hits
```{r}
full_results %>% filter(group == 'Gene', status == 'Hit', library == 'lncRNA') %>% pull(feature_id) %>% n_distinct()
```
### Differentiation hits
```{r}
differentiation_hits <- full_results %>% filter(group == 'Gene', status == 'Hit', assay == 'Differentiation') 
differentiation_hits %>% pull(feature_id) %>% n_distinct()
```
Coding differentiation hits
```{r}
differentiation_hits %>% filter(library == 'Coding') %>% pull(feature_id) %>% n_distinct()
```
lncRNA differentiation hits
```{r}
differentiation_hits %>% filter(library == 'lncRNA') %>% pull(feature_id) %>% n_distinct()
```
### Proliferation hits
```{r}
proliferation_hits <- full_results %>% filter(group == 'Gene', status == 'Hit', assay == 'Proliferation') 
proliferation_hits %>% pull(feature_id) %>% n_distinct()
```
Coding proliferation hits
```{r}
proliferation_hits %>% filter(library == 'Coding') %>% pull(feature_id) %>% n_distinct()
```
lncRNA proliferation hits
```{r}
proliferation_hits %>% filter(library == 'lncRNA') %>% pull(feature_id) %>% n_distinct()
```
### Dual hits
All
```{r}
dual_hits <- full_results %>% filter(group == 'Gene', status == 'Hit') %>% add_count(feature_id) %>% filter(n > 1)

dual_hits %>% pull(feature_id) %>%  n_distinct()
```
Coding
```{r}
dual_hits %>% filter(library == 'Coding') %>% pull(feature_id) %>% n_distinct()
```
lncRNA
```{r}
dual_hits %>% filter(library == 'lncRNA') %>%  pull(feature_id) %>% n_distinct()
```

## Mark primary phenotypes
Primary phenotype for hits is differentiation, proliferation, or dual.
Non-hits = Non-hit
```{r}
full_results_marked <- full_results %>% 
  mutate(primary_pheno = case_when(
    feature_id %in% dual_hits$feature_id ~ 'Dual',
    feature_id %in% differentiation_hits$feature_id ~ 'Differentiation',
    feature_id %in% proliferation_hits$feature_id ~ 'Proliferation',
    TRUE ~ 'None'),
    direction = ifelse(avg_pheno >= 0, 'Positive', 'Negative')) %>% 
  select(feature_id,
         gene_name,
         library,
         group,
         assay,
         pheno = avg_pheno,
         pval = avg_mw,
         score,
         threshold,
         control_sd,
         direction,
         status,
         hit_direction,
         primary_pheno,
         legend,
         experiment)

full_results_marked
```

## Export
```{r}
full_results_marked %>% write_tsv(file.path(analysis_dir, 'screen_results.tsv'))
results %>% write_rds(file.path(analysis_dir, 'results.rds'), compress = 'gz')
```

## Session info
```{r}
sessionInfo()
```


