---
title: "Permutation testing of phenotypes"
author: "David Wu"
output: html_notebook
---

## Setup 
Set working directory to project directory
```{r setup}
require(knitr)
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

### Optionally convert this Rmd to R script 
```{r eval=FALSE}
current_file <- rstudioapi::getActiveDocumentContext()$path
output_file <- stringr::str_replace(current_file, '.Rmd', '.R')
knitr::purl(current_file, output = output_file)
file.edit(output_file)
```
### Load libraries
```{r, message=FALSE}
library(tidyverse)
library(patchwork)
library(extrafont)
library(ggrepel)
library(ggforce)
library(eulerr)
library(parallel)
loadfonts()
```

### Load themes and functions
```{r}
source('analysis/universal/themes.R') # themes for all project components

source('analysis/universal/functions.R') # themes for all project components

source('analysis/screen/scripts/functions.R') # functions for this subcomponent

theme_set(theme_publication())
```

### Directories
```{r, message=FALSE}
analysis_dir <- 'analysis/screen/output/04_permutation_testing'
data_dir <- ''

dir.create(analysis_dir, showWarnings = FALSE)
dir.create(data_dir, showWarnings = FALSE)
```



### Import screen data
```{r}
screen_results <- read_tsv('analysis/screen/output/02_fdr/screen_results.tsv')
neighbor_integration <- read_tsv('analysis/screen/output/03_neighbors/neighbor_hits.tsv.gz')
```


### Permutation for negative vs positive skew
## Permute
```{r}
hits_table <- neighbor_integration %>% 
  filter(assay == 'Differentiation',
         full_status != 'Neighbor hit',
         (library == 'Coding' | locus_ambiguity != 'Ambiguous')) %>% 
  mutate(direction = ifelse(pheno > 0, 'Positive', 'Negative')) %>% 
  select(feature_id,
         library,
         assay,
         status,
         direction) 

hits_table
```

### Overall ratio

```{r}
hits_table %>% 
  filter(status == 'Hit') %>% 
  group_by(library, direction) %>% tally() %>% 
  pivot_wider(names_from = direction, values_from = n) %>% 
  mutate(ratio = round(Positive/Negative, 2),
         log2ratio = round(log2(ratio),2))
```

```{r}

permutations <- function(hits_table,
                         group1 = status,
                         group2 = direction) {
  
  
  permute_table <- hits_table %>% 
    dplyr::select(feature_id,
           library,
           {{group1}}, 
           {{group2}})
  
  colnames(permute_table) <- c('feature_id', 
                               'library',
                               'group1',
                               'group2')
  
  permute_table <- permute_table %>% 
    mutate(group1 = sample(group1),
           group2 = sample(group2))
  
  
  # tally
  permutation_summary <- permute_table %>% 
    select(-feature_id) %>% 
    group_by_all() %>% 
    tally() %>% 
    ungroup() %>% 
    mutate(description = paste(library, group1, group2, sep = '_')) %>% 
    select(description, n) %>% 
    pivot_wider(names_from = description, 
                values_from = n)
  

permutation_summary
}

permutations(hits_table)

```

### Trials for positive to negative ratio
#### All 
```{r eval=FALSE}
n_permutations <- 1e6
n_cores <- 70

tic()
permutations_direction <- mclapply(1:n_permutations, function(i) {
  
  permutation <- permutations(hits_table)
}, 
mc.cores = n_cores) %>% 
  bind_rows() %>% 
  mutate(Coding_Hit_ratio = Coding_Hit_Positive/Coding_Hit_Negative,
         lncRNA_Hit_ratio = lncRNA_Hit_Positive/lncRNA_Hit_Negative,
         `Coding_Non-hit_ratio` = `Coding_Non-hit_Positive`/`Coding_Non-hit_Negative`,
         `lncRNA_Non-hit_ratio` = `lncRNA_Non-hit_Positive`/`lncRNA_Non-hit_Negative`)
toc()

permutations_direction %>% write_tsv(file.path(analysis_dir, 'permutations_direction.tsv.gz'))

permutations_direction
```

#### Observed vs expected ratio
```{r}
observed_ratios <- hits_table %>% 
  filter(status == 'Hit') %>% 
  group_by(library, direction) %>% tally() %>% 
  pivot_wider(names_from = direction, values_from = n) %>% 
  mutate(ratio = round(Positive/Negative, 2),
         log2ratio = round(log2(ratio),2))

observed_coding_ratio <- observed_ratios %>% filter(library == 'Coding') %>% pull(ratio)
observed_lncrna_ratio <- observed_ratios %>% filter(library == 'lncRNA') %>% pull(ratio)

observed_coding_ratio

observed_lncrna_ratio
```

```{r}
coding_null_n <- permutations_direction %>% filter(log2(Coding_Hit_ratio) >= abs(log2(observed_coding_ratio)))

coding_p <- nrow(coding_null_n)/n_permutations

coding_p
```
```{r}
lncrna_null_n <- permutations_direction %>% filter(log2(lncRNA_Hit_ratio) >= abs(log2(observed_lncrna_ratio)))

lncrna_p <- nrow(lncrna_null_n)/n_permutations

lncrna_p
```



## Permutation for differentiation vs proliferation phenotype
### Permute
```{r}
hits_table <- neighbor_integration %>% 
  filter((library == 'Coding' | locus_ambiguity != 'Ambiguous')) %>% 
  mutate(status = ifelse(full_status == 'Neighbor hit', 'Non-hit', full_status),) %>% 
  select(feature_id,
         library,
         assay,
         status) %>% 
  pivot_wider(names_from = assay, values_from = status)

hits_table
```

```{r}
permute_table <- hits_table %>% 
  mutate(Differentiation = sample(Differentiation),
         Proliferation = sample(Proliferation))

permute_table

# tally
permutation_summary <- permute_table %>% 
  select(library, Differentiation, Proliferation) %>% 
  group_by(library, Differentiation, Proliferation) %>% 
  tally() %>% 
  mutate(primary_pheno = case_when(
    Differentiation == 'Hit' & Proliferation == 'Hit' ~ 'Dual',
    Differentiation == 'Hit' ~ 'Differentiation',
    Proliferation == 'Hit' ~ 'Proliferation',
    TRUE ~ 'Non-hit'),
    description = paste0(library, '_', primary_pheno)) %>% 
  ungroup() %>% 
  select(description, n) %>% 
  pivot_wider(names_from =  description,
              values_from = n) %>% 
  mutate(Coding_Hits = Coding_Differentiation + Coding_Proliferation,
         Coding_DP_Ratio = Coding_Differentiation/Coding_Proliferation,
         Coding_Dual_Ratio = Coding_Dual/Coding_Hits,
         lncRNA_Hits = lncRNA_Differentiation + lncRNA_Proliferation,
         lncRNA_DP_Ratio = lncRNA_Differentiation/lncRNA_Proliferation,
         lncRNA_Dual_Ratio = lncRNA_Dual/lncRNA_Hits,
         Coding_lncRNA_Ratio = Coding_DP_Ratio/lncRNA_DP_Ratio)

permutation_summary
```

### Trials for diff and prolif ratio
```{r}
tic()
permutations_phenotype <- mclapply(1:n_permutations, 
                                   mc.cores = n_cores,
                                   mc.preschedule = TRUE,
                                   function(i) {
                                     permute_table <- hits_table %>% 
                                       mutate(Differentiation = sample(Differentiation),
                                              Proliferation = sample(Proliferation))
                                     
                                     # tally
                                     permutation_summary <- permute_table %>% 
                                       select(library, Differentiation, Proliferation) %>% 
                                       group_by(library, Differentiation, Proliferation) %>% 
                                       tally() %>% 
                                       mutate(primary_pheno = case_when(
                                         Differentiation == 'Hit' & Proliferation == 'Hit' ~ 'Dual',
                                         Differentiation == 'Hit' ~ 'Differentiation',
                                         Proliferation == 'Hit' ~ 'Proliferation',
                                         TRUE ~ 'Non-hit'),
                                         description = paste0(library, '_', primary_pheno)) %>% 
                                       ungroup() %>% 
                                       select(description, n) %>% 
                                       pivot_wider(names_from =  description,
                                                   values_from = n)
                                     
                                     permutation_summary %>% 
                                       mutate('Trial' = i) %>% 
                                       select(Trial, everything())
                                   }) %>% 
  bind_rows() %>% 
  mutate(Coding_Hits = Coding_Differentiation + Coding_Proliferation,
         Coding_DP_Ratio = Coding_Differentiation/Coding_Proliferation,
         Coding_Dual_Ratio = Coding_Dual/Coding_Hits,
         lncRNA_Hits = lncRNA_Differentiation + lncRNA_Proliferation,
         lncRNA_DP_Ratio = lncRNA_Differentiation/lncRNA_Proliferation,
         lncRNA_Dual_Ratio = lncRNA_Dual/lncRNA_Hits,
         lncRNA_Coding_Ratio = lncRNA_DP_Ratio/Coding_DP_Ratio) %>% 
  mutate_if(is.double, round, 3)

permutations_phenotype %>% write_tsv(file.path(analysis_dir, 'permutations_phenotype.tsv.gz'))
toc()

permutations_phenotype
```


### Observed ratios

```{r}
observed_ratios <- hits_table %>% 
  filter(!(Differentiation == 'Hit' & Proliferation == 'Hit')) %>% 
  select(-feature_id) %>% 
  pivot_longer(cols = -library,
               names_to = 'assay',
               values_to = 'status') %>% 
  group_by_all() %>% 
  tally() %>% 
  filter(status == 'Hit') %>% 
  pivot_wider(names_from = assay,
              values_from = n) %>% 
  mutate(ratio = Differentiation/Proliferation)

observed_ratios
  
```
```{r}
observed_coding_ratio <- observed_ratios %>% filter(library == 'Coding') %>% pull(ratio)
observed_lncrna_ratio <- observed_ratios %>% filter(library == 'lncRNA') %>% pull(ratio)

observed_coding_ratio

observed_lncrna_ratio
```

```{r}
coding_null_n <- permutations_phenotype %>% filter(log2(Coding_DP_Ratio) >= abs(log2(observed_coding_ratio)))

coding_p <- nrow(coding_null_n)/n_permutations

coding_p
```
```{r}
lncrna_null_n <- permutations_phenotype %>% filter(log2(lncRNA_DP_Ratio) >= abs(log2(observed_lncrna_ratio)))

lncrna_p <- nrow(lncrna_null_n)/n_permutations

lncrna_p
```


### Plot
```{r}
permutations %>% 
  ggplot(aes(x = Coding_Dual_Ratio)) +
  geom_histogram(bins = bins, alpha = alpha) +
  labs(subtitle = 'Coding Hit Permutations',
       x = 'Differentiation:Proliferation Hit Ratio',
       y = 'Frequency')

lncrna <- permutations %>% 
  ggplot(aes(x = lncRNA_DP_Ratio)) +
  geom_histogram(bins = bins, alpha = alpha) +
  xlim(xlim) + 
  labs(subtitle = 'lncRNA Hit Permutations',
       x = 'Differentiation:Proliferation Hit Ratio',
       y = 'Frequency')
```

## Session info
```{r}
sessionInfo()
```

