---
title: "STRINGdb Protein Interaction Analysis"
author: "David Wu"
output: html_notebook
---
## Purpose
Analyze protein-protein interactions of hits and compare to random sets of genes

## Setup 
Set working directory to project directory
```{r setup}
require(knitr)
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

### Optionally convert this Rmd to R script 
```{r eval=FALSE}
current_file <- rstudioapi::getActiveDocumentContext()$path
output_file <- stringr::str_replace(current_file, '.Rmd', '.R')
knitr::purl(current_file, output = output_file)
file.edit(output_file)
```

### Load custom themes and functions
```{r, message=FALSE}
source('analysis/universal/themes.R') # themes for all project components

source('analysis/universal/functions.R') # themes for all project components
theme_set(theme_publication())
```

### Directories
```{r, message=FALSE}
analysis_dir <- 'analysis/screen/output/06_string' # analysis file output directory

dir.create(analysis_dir, showWarnings = FALSE)

```

### Load libraries
```{r, message=FALSE}
library(tidyverse)
library(extrafont)
loadfonts()
```

### Import screen data
```{r}
master <- read_tsv('analysis/integration/output/00_integrate/master_minimal.tsv.gz')
master
```

```{r}
coding_subset <- master %>% 
  filter(library == 'Coding')
```


```{r}
universe <- coding_subset %>% 
  pull(feature_id) %>% 
  unique()
```

## Download and load human STRINGdb network
Network: https://stringdb-static.org/download/protein.links.v11.0/9606.protein.links.v11.0.txt.gz
Mapping: https://stringdb-static.org/download/protein.info.v11.0/9606.protein.info.v11.0.txt.gz
```{r}
mapping <- read_tsv('analysis/screen/input/9606.protein.info.v11.0.txt.gz')
string <- read_delim('analysis/screen/input/9606.protein.links.v11.0.txt.gz', delim = ' ')
```

```{r}
mapped_universe <- mapping %>% filter(preferred_name %in% universe)
```
```{r}
string %>% head()
```
## Threshold interaction analysis
```{r}
interaction_network <- function(genes,
                                string,
                                mapping,
                                threshold = 150) {
  
  # map to STRINGdb identifiers
  mapped_ids <- mapping %>% filter(preferred_name %in% genes) %>% pull(protein_external_id)
  
  # subset protein network and filter for unique combinations
  network <- string %>% 
    filter(combined_score >= threshold,
      protein1 %in% mapped_ids,
      protein2 %in% mapped_ids) %>% 
    mutate(combination = ifelse(protein1 > protein2, paste(protein1, protein2), paste(protein2, protein1))) %>% 
    dplyr::select(combination, combined_score) %>% 
    unique()
  
  network
}

```

```{r}
hits <- coding_subset %>% filter(primary_pheno %in% c('Differentiation', 'Dual')) %>% pull(feature_id)

hits_network <- hits %>% interaction_network(string = string, mapping = mapping)

hits_network
```
## Random genes
```{r}
sample_size <- length(hits)
ntrials <- 1e3
cores <- 30
thresholds <- c(150,
                400,
                700)

tic()
random_trials <- mclapply(1:ntrials, 
         mc.cores = cores,
         FUN = function(i) {
  random_genes <- coding_subset %>%
    sample_n(sample_size) %>% 
    pull(feature_id)
  random_network <- random_genes %>% interaction_network(string = string, mapping = mapping)
  
  threshold_count <- map(thresholds, function(j) {
    
    passing_threshold <- random_network %>% 
      filter(combined_score >= j) %>% 
      nrow()
    
    tibble('threshold' = j,
           'interactions' = passing_threshold)
    
  }) %>% 
    bind_rows() %>% 
    mutate(trial = i)
}) %>% bind_rows() %>% 
  arrange(threshold, trial)

random_trials %>% head()
toc()
```


```{r}
random_summary <- random_trials %>% 
  group_by(threshold) %>% 
  summarize(mean = mean(interactions),
            sd = sd(interactions),
            conf1 = t.test(interactions, conf.level = 0.95)$conf.int[1],
            conf2 = t.test(interactions, conf.level = 0.95)$conf.int[2])

random_summary


```
```{r}
 hits_threshold <- map(thresholds, function(j) {
    
    passing_threshold <- hits_network %>% 
      filter(combined_score >= j) %>% 
      nrow()
    
    tibble('threshold' = j,
           'interactions' = passing_threshold)
    
  }) %>% 
    bind_rows()

hits_threshold
```
```{r}
combined_counts <- bind_rows(random_summary %>% 
                               dplyr::select(threshold,
                                             interactions = mean,
                                             sd,
                                             conf1,
                                             conf2) %>% 
                               mutate(group = 'Random'),
                             hits_threshold %>% 
                               mutate(group = 'Hits',
                                      sd = 0,
                                      conf1 = interactions,
                                      conf2 = interactions)) %>% 
  mutate(confidence = case_when(
    threshold == 150 ~ 'Low',
    threshold == 400 ~ 'Medium',
    threshold == 700 ~ 'High',
  ))
combined_counts
```
```{r}
p <- combined_counts %>% 
  mutate(confidence = factor(confidence, levels = c('Low', 'Medium', 'High', 'Highest'))) %>% 
  ggplot(aes(x = interactions,
             y = confidence,
             fill = group)) +
  geom_bar(stat = 'identity',
           position = 'dodge',
           #fill = 'darkred',
           alpha = 0.8) +
  geom_errorbar(aes(xmin=interactions - sd, xmax=interactions + sd), 
                width=.2,
                position=position_dodge(.9)) +
labs(x = 'Unique protein-protein interactions',
     y = 'Interaction confidence threshold') +
  scale_fill_few() +
  theme(legend.position = 'right')



save_figure(plot = p, 
            filename = 'string', 
            directory = 'figures/screen',
            w = 5,
            h = 3)

p 
  
```


 

## Export
```{r}
random_trials %>% write_tsv(file.path(analysis_dir, 'random_trials.tsv.gz'))
```

## Session info
```{r}
sessionInfo()
```
