---
title: "Integrate datasets"
author: "David Wu"
output:
  html_document:
    df_print: paged
---
## Purpose
Integrate genomic datasets into a master table

## Setup 
### Working directory
#### Set working directory to project directory (ipsc_manuscript)
```{r setup}
require(knitr)
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) # /media/data4/shared/ipsc_manuscript
```

### Load libraries
```{r, message=FALSE}
library(tidyverse)
library(patchwork)
library(ggthemes)
library(extrafont)
```

### Load themes and functions
```{r}
source('analysis/universal/themes.R') # themes for all project components

source('analysis/universal/functions.R') # themes for all project components

theme_set(theme_publication())
```


### Directories
```{r}
analysis_dir <- 'analysis/integration/output/enhancer'

dir.create(analysis_dir, showWarnings = FALSE)

```

### Master table
```{r}
master <- read_tsv('analysis/integration/output/000_integrate/master_wide.tsv.gz')
```

### Add column for any FANTOM overlap
```{r}
master <- master %>% 
  mutate(FANTOM5_any = rowSums(across(contains('FANTOM'))),
         FANTOM5_any = ifelse(FANTOM5_any > 0, 1, 0))

master %>% 
  select(feature_id, FANTOM5_any, contains('FANTOM'))
```
```{r}
master %>% filter(library == 'Coding' | library == 'Non-ambiguous')
```


```{r}
variables <- str_subset(colnames(master), 'FANTOM5|yosef')
```

### Test for enrichment across different categories
```{r}
enrich <- map(variables, function(i) {
  map(c('Coding', 'lncRNA'), function(which_library) {
    map(c('Differentiation', 'Dual', 'Proliferation', 'All', 'Differentiation+Dual', 'Proliferation+Dual'), function(which_pheno) {
      
      enrich_subset <- master %>% 
        filter(neighbor_overall_status != 'Neighbor hit',
               library == which_library,
               library == 'Coding' | locus_ambiguity == 'Non-ambiguous') %>% 
        select(feature_id, primary_pheno, dataset = {{i}})
      
      # primary pheno
      if(which_pheno == 'All') {
        pheno_subset <- c('Differentiation', 'Dual', 'Proliferation')
      } else if (which_pheno == 'Differentiation+Dual') {
        pheno_subset <- c('Differentiation', 'Dual')
      } else if (which_pheno == 'Proliferation+Dual') {
        pheno_subset <- c('Proliferation', 'Dual')
      } else {
        pheno_subset <- which_pheno
      }
      
      hits <- enrich_subset %>% filter(primary_pheno %in% pheno_subset) %>% pull(feature_id)  
      overlapping_features <- enrich_subset %>% filter(dataset == 1) %>% pull(feature_id) 
      
      universe <- enrich_subset %>% pull(feature_id) %>% unique()
      
      
      
      #Test enrichment for hits
      enrichment_results <- enrichment_test(list1 = hits,
                                            list2 = overlapping_features,
                                            cat1 = c('Hit', 'Non-hit'),
                                            cat2 = c('Overlap', 'Non-overlap'),
                                            background = universe,
                                            print = FALSE)
      
      double_positive_features <- intersect(hits,
                                            overlapping_features)
      
      tibble(library = which_library,
             primary_pheno = which_pheno,
             dataset = i,
             odds = enrichment_results$fisher$estimate,
             conf1 = enrichment_results$fisher$conf.int[1],
             conf2 = enrichment_results$fisher$conf.int[2],
             pvalue = enrichment_results$fisher$p.value,
             n_overlaps = double_positive_features %>% n_distinct(),
             n_hits = length(hits),
             n_features = length(overlapping_features),
             f_hits = round(n_overlaps/n_hits, 3))
      
    }) %>% bind_rows()
  }) %>% bind_rows() 
}) %>% bind_rows() %>% 
  mutate(log2odds = log2(odds),
         padj = p.adjust(pvalue, method = 'BH')) %>% 
  arrange(-n_overlaps)

  
enrich
  

```
### Subset to compare
```{r}
enrich_to_plot <- enrich %>% 
  filter(primary_pheno == 'Differentiation+Dual',
         dataset == 'FANTOM5_any')

p <- enrich_to_plot %>% 
  ggplot(aes(x = library,
             y = odds)) +
  geom_bar(stat = 'identity',
           fill = 'grey90',
           color = 'grey50') + 
  geom_errorbar(aes(ymin = conf1,
                    ymax = conf2),
                width = 0.2) +
  ylim(c(0, 4.5)) + 
  geom_hline(yintercept = 1, linetype = 'dashed', color = 'grey20') +
  labs(x = '',
       y = 'Odds',
       title = 'Odds of overlapping FANTOM5 enhancer')

save_figure(plot = p,
            filename = 'fantom5_overlap_odds',
            directory = 'figures/integration',
            h = 5, 
            w = 3)

p


```

### Subset to compare
```{r}
enrich_to_plot <- enrich %>% 
  filter(primary_pheno == 'Differentiation+Dual',
         dataset == 'ext_FANTOM5_neural')

p <- enrich_to_plot %>% 
  ggplot(aes(x = library,
             y = odds)) +
  geom_bar(stat = 'identity',
           fill = 'grey90',
           color = 'grey50') + 
  geom_errorbar(aes(ymin = conf1,
                    ymax = conf2),
                width = 0.2) +
  #ylim(c(0, 4.5)) + 
  geom_hline(yintercept = 1, linetype = 'dashed', color = 'grey20') +
  labs(x = '',
       y = 'Odds',
       title = 'Odds of overlapping FANTOM5 neural enhancer')

save_figure(plot = p,
            filename = 'fantom5_neural_overlap_odds',
            directory = 'figures/integration',
            h = 5, 
            w = 3)

p

```
## MPRA
```{r}
enrich_to_plot <- enrich %>% 
  filter(primary_pheno == 'All',
           dataset == 'ext_yosef_mpra')

p <- enrich_to_plot %>% 
  ggplot(aes(x = library,
             y = odds)) +
  geom_bar(stat = 'identity',
           fill = 'grey90',
           color = 'grey50') + 
  geom_errorbar(aes(ymin = conf1,
                    ymax = conf2),
                width = 0.2) +
  #ylim(c(0, 9)) + 
  geom_hline(yintercept = 1, linetype = 'dashed', color = 'grey20') +
  labs(x = '',
       y = 'Odds',
       title = 'Odds of overlapping FANTOM5 neural enhancer')

save_figure(plot = p,
            filename = 'yosef_overlap_odds',
            directory = 'figures/integration',
            h = 5, 
            w = 3)

p
```

### Individual
```{r}
which_library <-  'lncRNA'
which_pheno <- c('Differentiation','Dual')
which_dataset <- 'ext_FANTOM5_robust_enhancers'

enrich_subset <- master %>% 
  filter(neighbor_overall_status != 'Neighbor hit',
         library == 'Coding' | locus_ambiguity == 'Non-ambiguous',
         library == which_library) %>% 
  select(feature_id, primary_pheno, dataset = all_of(which_dataset))

hits <- enrich_subset %>% filter(primary_pheno %in% which_pheno) %>% pull(feature_id)
overlapping_features <- enrich_subset %>% filter(dataset == 1) %>% pull(feature_id) 

universe <- enrich_subset %>% pull(feature_id) %>% unique()



#Test enrichment for hits
enrichment_results <- enrichment_test(list1 = hits,
                                      list2 = overlapping_features,
                                      cat1 = c('Hit', 'Non-hit'),
                                            cat2 = c('Overlap', 'Non-overlap'),
                                            background = universe,
                                            print = TRUE)
```


How many hits intersect with FANTOM enhancers?
```{r}
master %>% 
  filter(FANTOM5_any == 1, neighbor_overall_status != 'Neighbor hit', primary_pheno !='None', library == 'Coding' | locus_ambiguity == 'Non-ambiguous') %>% 
  group_by(library) %>% tally()
```
How many are FANTOM enhancers from neural tissues?
```{r}
master %>% 
  filter(ext_FANTOM5_neural == 1, neighbor_overall_status != 'Neighbor hit', primary_pheno !='None') %>% 
  group_by(library) %>% tally()
```

How many hits intersect with Yosef neural induction cis-regulatory regions?
```{r}
master %>% 
  filter(ext_yosef_mpra == 1, neighbor_overall_status != 'Neighbor hit', primary_pheno != 'None') %>% 
  group_by(library) %>% tally()
```
```{r}
master %>% 
  filter(ext_yosef_mpra == 1, neighbor_overall_status != 'Neighbor hit', primary_pheno == 'None') %>% 
  group_by(library) %>% tally()
```

How many hits intersect with Yosef or FANTOM?
```{r}
master %>% 
  filter(ext_yosef_mpra == 1 | FANTOM5_any == 1, neighbor_overall_status != 'Neighbor hit', primary_pheno !='None') %>% 
  group_by(library) %>% tally()
```
```{r}
master %>% filter(enhancer == 'Non-enhancer', neighbor_overall_status == 'Non-neighbor') %>% group_by(library) %>% tally()
```

```{r}
enrich %>% 
  filter(primary_pheno %in% c('None', 'Differentiation', 'Proliferation', 'Dual'),
         dataset != 'FANTOM5_any') %>%  # equivalent to fantom5 robust
  mutate(padj = p.adjust(pvalue)) %>% 
  filter(dataset %in% c('ext_FANTOM5_robust_enhancers',
                        'ext_FANTOM5_neural',
                        'ext_yosef_mpra')) 
```
```{r}
enrich %>% 
  filter(primary_pheno %in% c('Differentiation+Dual')) %>%  
  filter(dataset %in% c('ext_FANTOM5_robust_enhancers',
                        'ext_FANTOM5_neural',
                        'ext_yosef_mpra')) %>% 
  mutate(padj = p.adjust(pvalue)) %>% 
  arrange(dataset, library) %>% 
  select(-primary_pheno) 

```

## Percentage table
```{r}
data_input <- master %>% 
  filter(diff_full_status != 'Neighbor hit',
         library == 'Coding' | locus_ambiguity == 'Non-ambiguous') %>% 
  mutate(status = diff_full_status)
  
group_totals <- 
  data_input %>% 
  group_by(library, status) %>% 
  tally(name = 'total')

group_totals
```

```{r}
fantom_any_count <- 
  data_input %>% 
  filter(FANTOM5_any == 1) %>% 
  group_by(library, status) %>% 
  tally() %>% 
  mutate(dataset = 'FANTOM Any') 

fantom_any_count 
  
```

```{r}
fantom_neural_count <- 
  data_input %>% 
  filter(ext_FANTOM5_neural == 1) %>% 
  group_by(library, status) %>% 
  tally() %>% 
  mutate(dataset = 'FANTOM Neural')

fantom_neural_count 
```

```{r}
mpra_count <- 
  data_input %>% 
  filter(ext_yosef_mpra == 1) %>% 
  group_by(library, status) %>% 
  tally() %>% 
  mutate(dataset = 'MPRA')

mpra_count
```

```{r}
overlap_percentages <- 
  bind_rows(list(fantom_any_count,
                 fantom_neural_count,
                 mpra_count)) %>% 
  left_join(group_totals) %>% 
  mutate(percentage = n / total * 100,
         status = factor(status, levels = c('Non-hit', 'Hit')))

overlap_percentages
```

```{r}
venn_colors <- c('#F29191', # pink
                 '#F24141', # red
                 '#F2B035', # orange
                 '#166B8C', # blue
                 '#26A699' # teal
                 )

ylim <- c(0, 20)
height <- 4
width <- 3
```

FANTOM5 any
```{r}
p <- overlap_percentages %>% 
  filter(dataset == 'FANTOM Any') %>% 
  ggplot(aes(x = library, 
             y = percentage,
             fill = status)) +
  geom_bar(stat = 'identity',
           position = 'dodge',
           alpha = 0.7) +
  ylim(ylim) +
  scale_fill_manual(values = c('grey70', venn_colors[4]))

save_figure(plot = p,
            filename = 'percent_fantom5_any',
            directory = 'figures/integration',
            h = height, 
            w = width)

p
```

FANTOM5 neural
```{r}
p <- overlap_percentages %>% 
  filter(dataset == 'FANTOM Neural') %>% 
  ggplot(aes(x = library, 
             y = percentage,
             fill = status)) +
  geom_bar(stat = 'identity',
           position = 'dodge',
           alpha = 0.7) +
  ylim(ylim) +
  scale_fill_manual(values = c('grey70', venn_colors[5]))

save_figure(plot = p,
            filename = 'percent_fantom5_neural',
            directory = 'figures/integration',
            h = height, 
            w = width)

p
```
MPRA
```{r}
p <- overlap_percentages %>% 
  filter(dataset == 'MPRA') %>% 
  ggplot(aes(x = library, 
             y = percentage,
             fill = status)) +
  geom_bar(stat = 'identity',
           position = 'dodge',
           alpha = 0.7) +
  ylim(ylim) +
  scale_fill_manual(values = c('grey70', venn_colors[3]))

save_figure(plot = p,
            filename = 'percent_mpra',
            directory = 'figures/integration',
            h = height, 
            w = width)

p
```
## Session info
```{r}
sessionInfo()
```
