---
title: 'Filter sgRNAs '
author: "David Wu"
output:
  html_document:
    df_print: paged
---
## Purpose
Analyze and filter sgRNA calls from Joseph Replogle

## Setup 
Set working directory to project directory
```{r setup}
require(knitr)
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

### Load custom themes and functions
```{r, message=FALSE}
source('analysis/universal/themes.R') # themes for all project components

source('analysis/universal/functions.R') # themes for all project components
theme_set(theme_publication())

source('analysis/perturbseq/scripts/functions.R')  # functions for this component; may be empty
```

### Directories
```{r, message=FALSE}
analysis_dir <- 'analysis/perturbseq/output/004_sgrna_filtering' # analysis file output directory
data_dir <- 'data/perturbseq/derived/seurat' # data file output directory

dir.create(analysis_dir, showWarnings = FALSE)
dir.create(data_dir, showWarnings = FALSE)
```

### Load libraries
```{r, message=FALSE}
library(tidyverse)
```

### Feature reference
```{r}
features <- read_csv('data/perturbseq/derived/pipestance/features.csv')
features
```

## Results
Joseph ran his pipeline and generated plots and also the final gRNA calls. 
Read data in here

```{r}
sgrna_calls <-
  read_csv('analysis/perturbseq/input/guide_calls/diff_cell_identities.csv') %>% mutate(condition = 'diff') %>%  # add condition
  mutate(sample = paste(condition, gemgroup, sep = '_'), # add sample id and fix numbering
         cell_barcode = str_replace(cell_barcode, '[:digit:]', '1'),
         unique_cell_id = paste(sample, cell_barcode, sep = '_'),
         combination = paste(unique_cell_id, guide_identity, sep = ':')) %>% 
  unique() %>% 
  add_count(unique_cell_id, name = 'num_features') # tally sgRNAs per cell

sgrna_calls
```

### Identify multiplets

```{r}
sgrna_calls %>% 
  ggplot(aes(x = num_features)) + 
  geom_histogram(binwidth = 1)
```

Rank sgRNAs and z-score
```{r}
scale_this <- function(x) as.vector(scale(x))

sgrna_augmented <- sgrna_calls %>% 
  group_by(guide_identity) %>% 
  mutate(z_score = scale_this(log10(UMI_count))) %>% # add z-scole per guide after log transformation
  ungroup() %>% 
  group_by(unique_cell_id) %>% 
  arrange(-UMI_count) %>% 
  mutate(rank = paste0('rank', 1:n()),
         sgrna_count = paste(guide_identity, UMI_count, round(z_score, 3), sep = ':')) %>% 
  ungroup() %>% 
  select(cell_barcode, guide_identity, combination, read_count, UMI_count, z_score, coverage, condition, sample, num_features, unique_cell_id, rank, sgrna_count)

sgrna_augmented
```



Plot UMI and read count 
```{r fig.width = 10, fig.height = 6}
map(conditions, function(i) {
  sgrna_augmented %>% 
    filter(condition == i) %>% 
    scatter_stats(x = 'read_count',
                  y = 'UMI_count',
                  log_scale = TRUE,
                  point_alpha = 0.1,
                  title = i)
}) %>% wrap_plots()
```
```{r fig.width = 10, fig.height = 8}
map(conditions, function(i) {
  sgrna_augmented %>% 
    filter(condition == i) %>% 
    scatter_stats(x = 'read_count',
                  y = 'coverage',
                  log_scale = TRUE,
                  point_alpha = 0.1,
                  title = i,
                  coord_equal = F) 
}) %>% wrap_plots()
```
Coverage
```{r}
sgrna_augmented %>% 
  ggplot(aes(x = coverage)) +
  geom_histogram(bins = 100) +
  facet_wrap(~condition) +
  scale_x_log10()
```

Pull singlet unique ids
```{r}
singlet_ids <- sgrna_calls %>% filter(num_features == 1) %>% pull(unique_cell_id)
singlet_ids %>% n_distinct()
```

Plot nUMI for rank 1 vs other sgRNA
```{r fig.width = 10, fig.height = 10}

  singlet <- sgrna_augmented %>% 
    filter(rank == 'rank1') %>% 
    scatter_stats(x = 'read_count',
                  y = 'UMI_count',
                  log_scale = TRUE,
                  point_alpha = 0.1)
  
  multiplet <- sgrna_augmented %>% 
    filter(rank != 'rank1') %>% 
    scatter_stats(x = 'read_count',
                  y = 'UMI_count',
                  log_scale = TRUE,
                  point_alpha = 0.1)
  
  list(singlet, multiplet) %>% wrap_plots(ncol = 1)
```
Coverage
```{r fig.width = 5, fig.height = 10}

  singlet <- sgrna_augmented %>% 
    filter(rank == 'rank1') %>% 
    scatter_stats(x = 'coverage',
                  y = 'UMI_count',
                  log_scale = TRUE,
                  point_alpha = 0.1,
                  coord_equal = FALSE)
  
  multiplet <- sgrna_augmented %>% 
    filter(rank != 'rank1') %>% 
    scatter_stats(x = 'coverage',
                  y = 'UMI_count',
                  log_scale = TRUE,
                  point_alpha = 0.1,
                  coord_equal = FALSE)
  
  list(singlet, multiplet) %>% wrap_plots(ncol = 1)
```


Mixture model 
```{r}
umi_mixture <- mixture_model(sgrna_augmented %>% pull(UMI_count) %>% log10(),
                             bins = 100)

umi_mixture$plot

```
Cells with less than 10 UMIs may be noisy
Median UMI per sgRNA
```{r}
median_umi <- sgrna_augmented %>% 
  group_by(guide_identity, condition) %>% 
  summarize(median_umi = median(UMI_count)) %>% arrange(median_umi)

median_umi
```

```{r}
median_umi %>% 
  ggplot(aes(x = median_umi)) +
  geom_histogram(bins = 100) +
  facet_wrap(~condition) +
  scale_x_log10()
```
For sgRNAs <10 median UMI, what is the distribution?

```{r}
low_umi_sgrna <- median_umi %>% filter(median_umi < 10) %>% pull(guide_identity) %>% unique()

sgrna_augmented %>% 
  filter(guide_identity %in% low_umi_sgrna) %>% 
  ggplot(aes(x = UMI_count)) + 
  geom_histogram(bins = 100) +
  facet_wrap(~condition) +
  scale_x_log10()
```
These may have been selectively lethal in iPSC, but distributed widely in the differentiation condition
```{r}
low_umi_plots <- map(low_umi_sgrna %>% set_names(.), function(i) {
  sgrna_augmented %>% 
  filter(guide_identity == i) %>% 
  ggplot(aes(x = UMI_count)) + 
  geom_histogram(bins = 30) +
  facet_wrap(~condition) +
  scale_x_log10() +
    ggtitle(i)
})

low_umi_plots 
```

Filter out UMI < 10 downstream after correcting for doublets

### Doublet analysis
Pull doublet unique ids and check if some doublets are likely singlets
```{r}
doublet_ids <- sgrna_augmented %>% 
  filter(num_features == 2) %>% 
  pull(unique_cell_id)

doublets <- sgrna_augmented %>% 
  filter(unique_cell_id %in% doublet_ids) %>% 
  select(unique_cell_id, sgrna_count, rank) %>% 
  pivot_wider(names_from = rank,
              values_from = sgrna_count) %>% 
  separate(rank1, into = c('rank1_sgrna', 'rank1_umi', 'rank1_z'), sep = ':') %>% 
  separate(rank2, into = c('rank2_sgrna', 'rank2_umi', 'rank2_z'), sep = ':') %>% 
  mutate(rank1_target = str_sub(rank1_sgrna, end = -3),
         rank2_target = str_sub(rank2_sgrna, end = -3)) %>% 
  mutate(rank1_umi = as.integer(rank1_umi),
         rank2_umi = as.integer(rank2_umi),
         rank1_z = as.numeric(rank1_z),
         rank2_z = as.numeric(rank2_z),
         umi_difference = rank1_umi - rank2_umi,
         umi_ratio = rank1_umi/rank2_umi,
         umi_log10ratio = log10(umi_ratio),
         z_difference = rank1_z - rank2_z,
         z_ratio = rank1_z/rank2_z,
         sample = str_sub(unique_cell_id, end = 6),
         condition = str_sub(sample, end = 4))

doublets 
```

Identify doublets where the rank2 sgRNA is likely to be noise and cell is actually a singlet
```{r, fig.height=5, fig.width=5}
doublets %>% 
  scatter_stats(x = 'rank1_umi',
                y = 'rank2_umi',
                log_scale = TRUE, 
                pseudocount = 0,
                point_size = 0.5,
                point_alpha = 0.1)
```

Without controls, splitting by condition
```{r, fig.height=5, fig.width=10}
  doublets %>% 
    filter(rank1_target != 'sgNTC',
           rank2_target != 'sgNTC') %>% # filter out controls 
    scatter_stats(x = 'rank1_umi',
                  y = 'rank2_umi',
                  log_scale = TRUE, 
                  pseudocount = 0,
                  point_size = 0.5,
                  point_alpha = 0.1)
```
Plot z-scores
```{r, fig.height=5, fig.width=10}

  doublets %>% 
    filter(rank1_target != 'sgNTC',
           rank2_target != 'sgNTC') %>% # filter out controls 
    scatter_stats(x = 'rank1_z',
                  y = 'rank2_z',
                  log_scale = FALSE, 
                  pseudocount = 0,
                  point_size = 0.5,
                  point_alpha = 0.1)
```


Plot z-score differences between rank 1 and 2 
Appears to be Gaussian mixture
```{r}
doublets %>% 
  ggplot(aes(x = z_difference)) +
  geom_histogram(bins = 5000,
                 alpha = 0.8) +
  facet_wrap(~condition)
```
Mixture model on z-scores
```{r}
doublets_mixture <- mixture_model(doublets %>% pull(z_difference), xlab = 'Z-score Difference')
doublets_mixture$plot
```
```{r}
doublets_mixture$cutoff
```

```{r}
doublets_mixture$model$mu
```
Mixture on rank2 z-score
```{r}
rank2_mixture <- mixture_model(doublets %>% pull(rank2_z), xlab = 'Z-score Difference')
rank2_mixture$plot
```

```{r}
rank2_mixture$cutoff
```

```{r}
rank2_mixture$model$mu
```

Use a cutoff of a z-score difference and rank2 for potential singlets
```{r}
potential_singlets <- doublets %>% 
  filter(z_difference > max(doublets_mixture$model$mu),
         rank2_z < min(rank2_mixture$model$mu)) %>% 
  select(unique_cell_id, guide_identity = rank1_sgrna) %>% 
  mutate(combination = paste(unique_cell_id, guide_identity, sep = ':'))

potential_singlets
```

Filtered sgRNA table
```{r}
filtered_sgrnas <- sgrna_augmented %>% 
  filter(unique_cell_id %in% c(singlet_ids, doublet_ids)) %>% 
  select(unique_cell_id, condition, sample, cell_barcode, UMI_count, z_score, rank, guide_identity) %>% 
  left_join(potential_singlets %>% select(unique_cell_id, sgrna_call = guide_identity)) %>% 
  mutate(sgrna_call = ifelse(is.na(sgrna_call), guide_identity, sgrna_call),
         status = case_when(
           unique_cell_id %in% singlet_ids ~ 'Singlet',
           unique_cell_id %in% potential_singlets$unique_cell_id ~ 'Undetermined',
           unique_cell_id %in% doublet_ids ~ 'Doublet')) %>% 
  left_join(features %>% select(guide_identity = id,
                                guide_target = target_gene_id)) %>% 
  arrange(desc(status), unique_cell_id)

filtered_sgrnas 
```

```{r}
filtered_sgrnas %>% 
  select(unique_cell_id,
         status) %>% 
  unique() %>% 
  group_by(status) %>% 
  tally()
``` 

### Cells per sgRNA
```{r}
cells_per_sgrna <- filtered_sgrnas %>% 
  group_by(condition, sgrna_call) %>% 
  tally() %>% 
  arrange(n)

cells_per_sgrna
```
```{r}
cells_per_sgrna %>% 
  filter(str_detect(sgrna_call, 'NTC', negate = TRUE)) %>% 
  ggplot(aes(x = n)) +
  geom_histogram() + 
  facet_wrap(~condition) +
  scale_x_log10()
```

## Plot
### sgRNA UMIs per cell
```{r}
filtered_sgrnas %>% 
  ggplot(aes(x = sample, 
             y = UMI_count,
             color = sample)) + 
  geom_sina(method = 'density', alpha = 0.2, size = 0.1) + 
  scale_y_log10() + 
  ylab('sgRNA UMIs per cell') +
  annotation_logticks()
```
#### All counts
```{r}
filtered_sgrnas %>% 
  group_by(sample) %>% 
  summarize(mean = mean(UMI_count),
            sd = sd(UMI_count),
            median = median(UMI_count),
            iqr = IQR(UMI_count)
            ) %>% 
  mutate_if(is.numeric, round)
```

#### Singlets
```{r}
filtered_sgrnas %>% 
  filter(status == 'Singlet') %>% 
  group_by(sample) %>% 
  summarize(mean = mean(UMI_count),
            sd = sd(UMI_count),
            median = median(UMI_count),
            iqr = IQR(UMI_count)
            ) %>% 
  mutate_if(is.numeric, round)
```

### Cells per sgRNA
```{r}
cells_per_sgrna %>% 
  ungroup() %>% 
  filter(str_detect(sgrna_call, 'NTC', negate = TRUE)) %>% 
  ggplot(aes(x = condition,
             y = n,
             color = condition)) +
  geom_sina(method = 'density', alpha = 0.5, size = 1) +
  scale_color_few() +
  scale_y_log10()
```
```{r}
cells_per_sgrna %>% 
  group_by(condition) %>% 
  summarize(mean = mean(n),
            sd = sd(n),
            median = median(n),
            iqr = IQR(n)) %>% 
  mutate_if(is.numeric, round)
```

### Cells per guide target
```{r}
cells_per_target <- filtered_sgrnas %>% 
  filter(status == 'Singlet') %>% 
  group_by(condition, guide_target) %>% 
  tally() %>% 
  arrange(n)

cells_per_target %>% 
  filter(str_detect(guide_target, 'NTC', negate = TRUE)) %>% 
  ggplot(aes(x = condition,
             y = n,
             color = condition)) +
  geom_sina(method = 'density', alpha = 0.5, size = 1) +
  scale_color_few() +
  scale_y_log10()
```
```{r}
cells_per_target %>% 
  group_by(condition) %>% 
  summarize(mean = mean(n),
            sd = sd(n),
            median = median(n),
            iqr = IQR(n)) %>% 
  mutate_if(is.numeric, round)
```


## Export
```{r}
filtered_sgrnas %>% write_tsv(file.path(analysis_dir, 'filtered_sgrnas_replogle.tsv.gz'))
```

## Session info
```{r}
sessionInfo()
```
