---
title: "Load into Seurat"
author: "David Wu"
output: html_notebook
---
## Purpose
Generate Seurat objects from imported data and add metadata

## Setup 
Set working directory to project directory
```{r setup}
require(knitr)
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

### Optionally convert this Rmd to R script 
```{r eval=FALSE}
current_file <- rstudioapi::getActiveDocumentContext()$path
output_file <- stringr::str_replace(current_file, '.Rmd', '.R')
knitr::purl(current_file, output = output_file)
file.edit(output_file)
```
### Load custom themes and functions
```{r, message=FALSE}
source('analysis/universal/themes.R') # themes for all project components

source('analysis/universal/functions.R') # themes for all project components
theme_set(theme_publication())

source('analysis/perturbseq/scripts/functions.R')  # functions for this component; may be empty
```

### Directories
```{r, message=FALSE}
analysis_dir <- 'analysis/perturbseq/output/006_seurat' # analysis file output directory
data_dir <- 'data/perturbseq/derived/seurat' # data file output directory

dir.create(analysis_dir, showWarnings = FALSE)
dir.create(data_dir, showWarnings = FALSE)
```

### Load libraries
```{r, message=FALSE}
library(tidyverse)
library(Seurat)
library(parallel)
```

### Import feature reference
```{r}
features <- read_csv('data/perturbseq/derived/pipestance/features.csv')
features
```
### Feature calls
```{r}
filtered_sgrnas <- read_tsv('analysis/perturbseq/output/004_sgrna_filtering/filtered_sgrnas.tsv.gz') %>% 
  left_join(features %>% select(guide_identity = id,
                                guide_target = target_gene_id))
filtered_sgrnas
```

### sgRNA metadata to add
```{r}
metadata_table <- filtered_sgrnas %>% 
  select(cell_barcode,
         sample, 
         condition,
         UMI_count, 
         z_score, 
         guide_identity,
         guide_target,
         singlet_status) 

metadata_table
```

### Read in filtered matrices from 005_import
```{r}
tic()
unified_cellranger_matrices <- read_rds(file.path(data_dir, 'unified_cellranger_matrices.rds'))
unified_cellbender_matrices <- read_rds(file.path(data_dir, 'unified_cellbender_matrices.rds'))
toc()
```

## Unified
## Import data and load into Seurat
### Identify output files
```{r}
# Pull out directories of interest
pipestance_dir <- 'data/perturbseq/derived/pipestance'
pipestances <- list.dirs(pipestance_dir, full.names = T, recursive = F) %>% str_subset('diff')
sample_names <- basename(pipestances) %>% set_names(.)
names(pipestances) <- sample_names
pipestances
```

### Load into Seurat object
```{r}
tic() 
unified_cellranger <- map(sample_names, function(sample) {
  so <- CreateSeuratObject(unified_cellranger_matrices[[sample]]$gene, project = sample)
  so[['grna']] <- CreateAssayObject(counts = unified_cellranger_matrices[[sample]]$grna) # cellranger sgRNA counts
  so
  })
toc()
```

```{r}
unified_cellranger
```

## Add metadata
```{r}
tic()
unified_cellranger <- map(unified_cellranger, function(seurat) {
  barcodes <- colnames(seurat)
  metadata <- as.data.frame(metadata_table %>% 
                              filter(sample == seurat@project.name))
  rownames(metadata) <- metadata$cell_barcode
  metadata <- metadata[barcodes, -1]
  AddMetaData(seurat, metadata)
})
toc()
```

```{r}
unified_cellranger <- unified_cellranger %>% map(function(seurat) {PercentageFeatureSet(seurat, pattern = '^MT-', col.name = 'percent_mito')})
unified_cellranger <- unified_cellranger %>% map(function(seurat) {PercentageFeatureSet(seurat, pattern = '^RPS|^RPL|^MRPS|^MRPL', col.name = 'percent_ribo')})
```

### Filter out zeroes
```{r}
tic()
unified_cellranger <- map(unified_cellranger, function(seurat) {
  
  barcodes <- seurat@meta.data %>% 
    rownames_to_column() %>% 
    filter(nCount_RNA > 0,
           percent_mito > 0,
           percent_ribo > 0) %>% 
    pull(rowname)
  
  subset(seurat, cells = barcodes)
})
toc()
```

```{r}
unified_cellranger
```
## Export
```{r}
tic()

unified_cellranger %>% write_rds(file.path(data_dir, 'unified_cellranger.rds'), compress = 'gz')
toc()
```
```{r}
rm(unified_cellranger)
gc()
```

## Unified cellbender

```{r}
tic() 
unified_cellbender <- map(sample_names, function(sample) {
  so <- CreateSeuratObject(unified_cellbender_matrices[[sample]]$gene, project = sample)

  so
  })
toc()
```

### Add sgRNA metadata
```{r}
tic()
unified_cellbender <- map(unified_cellbender, function(seurat) {
  barcodes <- colnames(seurat)
  metadata <- as.data.frame(metadata_table %>% 
                              filter(sample == seurat@project.name))
  rownames(metadata) <- metadata$cell_barcode
  metadata <- metadata[barcodes, -1]
  AddMetaData(seurat, metadata)
})
toc()

```

### Add qc metrics (ribosomal and mitochondrial content)
```{r}
tic()

unified_cellbender <- unified_cellbender %>% map(function(seurat) {PercentageFeatureSet(seurat, pattern = '^MT-', col.name = 'percent_mito')})
unified_cellbender <- unified_cellbender %>% map(function(seurat) {PercentageFeatureSet(seurat, pattern = '^RPS|^RPL|^MRPS|^MRPL', col.name = 'percent_ribo')})

toc()
```

### Exclude zeros
```{r}
tic()
unified_cellbender <- map(unified_cellbender, function(seurat) {
  barcodes <- seurat@meta.data %>% 
    rownames_to_column() %>% 
    filter(nCount_RNA > 0,
           percent_mito > 0,
           percent_ribo > 0) %>% 
    pull(rowname)
  
  subset(seurat, cells = barcodes)
  
})
toc()


```
```{r}
unified_cellbender
```

## Export
```{r}
unified_cellbender %>% write_rds(file.path(data_dir, 'unified_cellbender.rds'), compress = 'gz')
```


## Session info
```{r}
sessionInfo()
```

