---
title: "QC analysis of Perturb-Seq data"
author: "David Wu"
output: html_notebook
---
## Purpose
Identify QC clusters and run pipeline to examine what cell type clusters they correspond to downstream

## Setup 
Set working directory to project directory
```{r setup}
require(knitr)
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

### Optionally convert this Rmd to R script 
```{r eval=FALSE}
current_file <- rstudioapi::getActiveDocumentContext()$path
output_file <- stringr::str_replace(current_file, '.Rmd', '.R')
knitr::purl(current_file, output = output_file)
file.edit(output_file)
```
### Load custom themes and functions
```{r, message=FALSE}
source('analysis/universal/themes.R') # themes for all project components

source('analysis/universal/functions.R') # themes for all project components
theme_set(theme_publication())

source('analysis/perturbseq/scripts/functions.R')  # functions for this component; may be empty
```

### Functions 
```{r}
tabulate_frequency <- function(metadata_table,
                               group1 = guide_target,
                               group2 = seurat_clusters,
                               rename = TRUE) {
  frequency_table <- 
      metadata_table %>% 
    mutate_all(as.factor) %>% 
    mutate_all(droplevels) %>% 
    dplyr::select({{group1}}, 
                  {{group2}}) 
  
  if(rename) {
    colnames(frequency_table) <- c('target', 'cluster')
  }

  frequency_table %>% table()
  
}

calculate_proportions <- function(frequency_table) {
  
  target_proportion <- prop.table(frequency_table, margin = 1) %>% 
    as.data.frame() %>% 
    as_tibble() %>% 
    mutate(target_proportion = round(Freq, 3)) %>% 
    dplyr::select(-Freq)
  
  cluster_proportion <- prop.table(frequency_table, margin = 2) %>% 
    as.data.frame() %>% 
    as_tibble() %>% 
    mutate(cluster_proportion = round(Freq, 3)) %>% 
    dplyr::select(-Freq)
  
  frequency_long <- frequency_table %>% 
    as.data.frame() %>% 
    as_tibble()
  
  colnames(frequency_long) <- c('target', 'cluster', 'n_overlap')
  
  cluster_totals <- colSums(frequency_table) %>% 
    as.data.frame() %>% 
    rownames_to_column()
  
  colnames(cluster_totals) <- c('cluster', 'n_cluster')
  
  target_totals <- rowSums(frequency_table) %>% 
    as.data.frame() %>% 
    rownames_to_column()
  colnames(target_totals) <- c('target', 'n_target')
  
  proportion_table <- inner_join(cluster_proportion,
                                 target_proportion) %>% 
    inner_join(frequency_long) %>% 
    inner_join(target_totals) %>% 
    inner_join(cluster_totals) %>% 
    mutate(max_proportion = ifelse(cluster_proportion > target_proportion, cluster_proportion, target_proportion),
           score = round(cluster_proportion * target_proportion, 3)) %>% 
    arrange(-max_proportion)
  
  proportion_table
  
}

proportion_histograms <- function(proportion_table) {
  list(proportion_table %>% ggplot(aes(x = cluster_proportion)) + geom_histogram() + scale_x_log10(),
       proportion_table %>% ggplot(aes(x = target_proportion)) + geom_histogram() + scale_x_log10(),
       proportion_table %>% ggplot(aes(x = score)) + geom_histogram() + scale_x_log10()) %>% 
    wrap_plots()
}

create_proportion_matrix <- function(frequency_table) {
  target_proportion <- prop.table(frequency_table, margin = 1)
  
  cluster_proportion <- prop.table(frequency_table, margin = 2)
  
  list('cluster' = as.matrix(cluster_proportion),
       'target' = as.matrix(target_proportion))
}

proportion_analysis <- function(metadata_table,
                               group1 = guide_target,
                               group2 = seurat_clusters,
                               rename = TRUE,
                               heatmap_names = FALSE) {
  
  condition <- metadata_table$condition %>% unique()
  object <- metadata_table$object %>% unique()
  
  freq_table <- metadata_table %>% 
    tabulate_frequency(group1 = {{group1}},
                       group2 = {{group2}},
                       rename = rename)
  
  prop_table <- freq_table %>% calculate_proportions()
  prop_matrix <- freq_table %>% create_proportion_matrix()
  prop_hist <- prop_table %>% proportion_histograms()
  
  cluster_heatmap <- pheatmap(t(prop_matrix$cluster), 
                              scale = 'column', 
                              clustering_distance_rows = 'correlation', 
                              clustering_distance_cols = 'correlation',
                              main = paste(condition, object, 'cluster'),
                              show_colnames = heatmap_names)
  
  target_heatmap <- pheatmap(t(prop_matrix$target), 
                             scale = 'column', 
                             clustering_distance_rows = 'correlation', 
                             clustering_distance_cols = 'correlation',
                             main = paste(condition, object, 'target'),
                             show_colnames = heatmap_names)
    
  heatmaps <- list('cluster' = cluster_heatmap,
                   'target' = target_heatmap)
  
  list('frequency' = freq_table,
       'proportions' = prop_table,
       'matrix' = prop_matrix,
       'histogram' = prop_hist,
       'heatmap' = heatmaps)
}
```
### Directories
```{r, message=FALSE}
analysis_dir <- 'analysis/perturbseq/output/007_qc_analysis' # analysis file output directory
data_dir <- 'data/perturbseq/derived/seurat' # data file output directory

dir.create(analysis_dir, showWarnings = FALSE)
dir.create(data_dir, showWarnings = FALSE)
```

### Load libraries
```{r, message=FALSE}
library(tidyverse)
library(Seurat)
library(plotly)
library(cluster)
library(pheatmap)
```

## cellranger
## Import Seurat object
```{r}
seurats <- read_rds('data/perturbseq/derived/seurat/unified_cellranger.rds')
```
```{r}
seurats
```


```{r}
# other cutoffs
umi_cutoff <- 2000
complexity_cutoff <- 1000
```

```{r fig.height=8, fig.width=16}
seurats %>% 
  map(seurat_statsplot, vline = umi_cutoff, hline = complexity_cutoff) %>% 
  wrap_plots(nrow = 2)
```

```{r fig.height=8, fig.width=16}
seurats %>% 
  map(seurat_statsplot, x = 'percent_mito', y = 'percent_ribo', logx = F, logy = F, median_lines = F) %>% 
  wrap_plots(nrow = 2)
```

```{r fig.height=8, fig.width=16}
seurats %>% 
  map(seurat_statsplot, x = 'percent_mito', y = 'nFeature_RNA', logx = F, logy = F, median_lines = F) %>% 
  wrap_plots(nrow = 2)
```

Perform k-medioid clustering to roughly partition cells based on QC metrics  
Set k = 3
```{r fig.height=8, fig.width=16}
tic()
seurats <- map(seurats, function(seurat) {
  cat(paste(seurat@project.name, '\n'))
  qc_metrics <- seurat@meta.data[,c(
    'nFeature_RNA', # just use nFeature since it is nearly redundant with nUMI
    'percent_ribo',
    'percent_mito'
  )] %>% 
    rownames_to_column() %>% 
    mutate_at(vars(contains('_')), log10) %>% # log transform 
    mutate_if(is.numeric, scale) %>% # scale for clustering
    column_to_rownames()
  
  # # kmeans
  clusters <- kmeans(qc_metrics, 3, nstart = 30)
  cluster_labels <- clusters$cluster
  
  # use consistent coloring
  qc_metrics$cluster <- cluster_labels

  
  cluster_tabulate <- cluster_labels %>% table() %>% sort(decreasing = T)
  
  # order clusters by mitochondrial percentage
  qc_summary <- qc_metrics %>% group_by(cluster) %>% summarize_all(median)
  cluster_order <- qc_summary %>% arrange(percent_mito) %>% pull(cluster)
  
  cluster_labels[cluster_labels == cluster_order[1]] <- 'QC1'
  cluster_labels[cluster_labels == cluster_order[2]] <- 'QC2'
  cluster_labels[cluster_labels == cluster_order[3]] <- 'QC3'

seurat <- AddMetaData(seurat, data.frame('qc_cluster' = as.factor(cluster_labels)))

})
toc()

map(seurats, function(seurat) {
  seurat@meta.data %>% 
  ggplot(aes(x = percent_ribo,
             y = log10(nCount_RNA),
             color = qc_cluster)) + 
  geom_point(alpha = 0.1, size = 0.5) +
  scale_color_few() +
    ggtitle(seurat@project.name) + 
    theme(legend.position = 'bottom')
}) %>% wrap_plots(nrow = 2)
```
```{r fig.height=8, fig.width=16}
map(seurats, function(seurat) {
  seurat@meta.data %>% 
  ggplot(aes(x = percent_mito,
             y = log10(nFeature_RNA),
             color = qc_cluster)) + 
  geom_point(alpha = 0.1, size = 0.5) +
  scale_color_few() +
    ggtitle(seurat@project.name) + 
    theme(legend.position = 'bottom')
}) %>% wrap_plots(nrow = 2)
```
```{r fig.height=8, fig.width=16}
map(seurats, function(seurat) {
  seurat@meta.data %>% 
  ggplot(aes(x = percent_ribo,
             y = log10(nFeature_RNA),
             color = qc_cluster)) + 
  geom_point(alpha = 0.5, size = 0.25) +
  scale_color_few() +
    ggtitle(seurat@project.name) + 
    theme(legend.position = 'bottom')
}) %>% wrap_plots(nrow = 2)
```


### Facet by QC cluster
```{r fig.height=12, fig.width=16}
map(seurats, function(seurat) {
  seurat@meta.data %>% 
    ggplot(aes(x = percent_ribo,
               y = log10(nCount_RNA),
               color = qc_cluster)) + 
    geom_point(alpha = 0.8, size = 0.25) +
    scale_color_few() +
    facet_wrap(~qc_cluster, nrow = 2) + 
    ggtitle(seurat@project.name) + 
    theme(legend.position = 'bottom')
}) %>% wrap_plots(nrow = 2)
```
```{r}
seurat <- seurats$diff_1
    
plot_ly(seurat@meta.data,
        x = ~log10(nFeature_RNA),
        y = ~log10(percent_mito),
        z = ~log10(percent_ribo),
        color = ~qc_cluster,
        marker = list(size = 5))
```


```{r}
seurats %>% write_rds(file.path(data_dir, 'unified_cellranger_qc.rds'))
```


## Pipeline
Proceed with basic pipeline to understand how much QC affects downstream analyses
### Set up
```{r}
conditions <- list('diff' = seurats)

dims_to_integrate <- 1:30
```

### Integration 
Use reference-based since all samples are similar
Use largest sample as reference (for each condition)
```{r}
tic()
merged <- map(conditions, function(i) {
  
  seurat_list <- i %>% 
    map(NormalizeData) %>% 
    map(FindVariableFeatures, 
        selection.method = 'vst', 
        nfeatures = 1000)
  
  largest_sample <- map(seurat_list, ncol) %>% unlist() %>% sort(decreasing = TRUE) %>% names() %>% .[1]
  reference_sample <- which(names(seurat_list) == largest_sample)
  
  integration_features <- SelectIntegrationFeatures(object.list = seurat_list,
                                                    nfeatures = 1000)
  
  anchors <- FindIntegrationAnchors(object.list = seurat_list,
                                    anchor.features = integration_features,
                                    dims = dims_to_integrate,
                                    reference = reference_sample)
  
  integrated <- IntegrateData(anchorset = anchors,
                              dims = dims_to_integrate)
  
  integrated
})
toc()
```


### Dimensionality reduction
```{r}
calculate_dims <- 50

merged <- merged %>% 
  map(ScaleData) %>% 
  map(RunPCA, npcs = calculate_dims) 

merged %>% map(ElbowPlot, ndims = calculate_dims)
```



```{r}
significant_dims <- 1:30
k <- 30
neighbors <- 30
min_dist <- 0.1
n_epochs <- 200
umap_seed <- 20191214
threads <- 24

tic()
merged <- merged %>% map(FindNeighbors, dims = significant_dims, k.param = neighbors) # for SNN, use lower K 
toc()

tic()
merged <- merged %>% map(FindClusters, algorithm = 1, resolution = 1.6) # use high resolution to find subclusters
toc()
```


```{r}
tic()
merged <- merged %>% map(RunUMAP, 
                         dims = significant_dims, 
                         reduction = 'pca', 
                         reduction.name = 'umap',
                         n.components = 2, 
                         n.neighbors = k, 
                         min.dist = min_dist,
                         n.epochs = n_epochs,
                         n_threads = 24,
                         seed.use = umap_seed)
toc()

merged %>% map(seurat_plot, reduction = 'umap', feature = 'ident', do.label = T) %>% wrap_plots()
```
```{r}
tsne_seed <- umap_seed
theta <- 0.3
iter <- 1000
tic()
merged <- merged %>% map(RunTSNE, 
                         dims = significant_dims, 
                         reduction = 'pca', 
                         tsne.method = 'FIt-SNE', 
                         fast_tsne_path = '/home/software/FIt-SNE/bin/fast_tsne', 
                         max_iter = iter, 
                         nthreads = threads,
                         theta = theta, 
                         perplexity = k,
                         seed.use = tsne_seed);
toc()

merged %>% map(seurat_plot, reduction = 'tsne', 
               feature = 'ident', 
               size = 0.5, alpha = 0.5, color_package = 'hues', color_scale = 'pimp', do.label = T,
               show.legend = F) %>% wrap_plots()
```

Check original QC clusters
```{r}
merged %>% map(seurat_plot, reduction = 'umap', feature = 'qc_cluster', size = 0.3, alpha = 0.1, color_package = 'carto', color_scale = 'Prism', do.label = T) %>% wrap_plots()
```
```{r}
merged %>% map(seurat_plot, reduction = 'tsne', feature = 'qc_cluster', size = 0.3, alpha = 0.3, color_package = 'carto', color_scale = 'Prism', do.label = T) %>% wrap_plots()
```

Plot mitochondrial content
```{r}
merged %>% map(seurat_feature, reduction = 'umap', gene = 'percent_mito', size = 0.3, alpha = 0.3, colors = c('grey80', 'grey80', 'dodgerblue4')) %>% wrap_plots()
```

```{r}
merged %>% map(seurat_feature, reduction = 'umap', gene = 'percent_ribo', size = 0.3, alpha = 0.3, colors = c('grey80', 'grey80', 'dodgerblue4')) %>% wrap_plots()
```

```{r}
merged %>% map(seurat_feature, reduction = 'umap', gene = 'nFeature_RNA', size = 0.3, alpha = 0.3, colors = c('grey80', 'grey80', 'dodgerblue4')) %>% wrap_plots()
```
## Cluster analysis
### Are specific sgRNAs enriched in any of the low QC clusters?
```{r}
groups <- c('diff') %>% set_names(.)

cluster_table <- map(groups, function(i) {
  merged[[i]]@meta.data %>% 
    rownames_to_column('barcode')
})
  
cluster_table
```

### Frequency analysis
QC frequency
```{r}
qc_frequency <- map(groups, function(i) {
  freq <- cluster_table[[i]] %>% 
    tabulate_frequency(group1 = qc_cluster,
                       group2 = seurat_clusters)
  
  prop <- calculate_proportions(freq) %>% mutate(condition = i)
})

qc_frequency
```

Which cell clusters overlap with low QC clusters?
```{r}
qc_frequency$diff %>% 
  filter(target != 'QC2',
         cluster_proportion > 0.2) 
```

```{r}
low_qc_clusters <- map(groups, function(i) {
  qc_frequency[[i]] %>% 
  filter(target == 'QC3',
         cluster_proportion > 0.2) %>% 
  pull(cluster)
})

low_qc_clusters
```

Are any targets enriched?
```{r}
target_frequency <- map(groups, function(i) {
  freq <- cluster_table[[i]] %>% 
    tabulate_frequency(group1 = guide_target,
                       group2 = seurat_clusters)
  
  prop <- calculate_proportions(freq) %>% mutate(condition = i)
})

target_frequency
```

```{r}
target_frequency$diff %>% 
  filter(cluster %in% low_qc_clusters$diff)
```

For targets found in poor QC clusters, some appear to be enriched for certain sgRNAs. 
Therefore, as a first pass, keep all filtered singlets for downstream analysis. Maintain QC cluster metadata in seurat object for future filtering if desired.

If further filtering is desired, identify barcodes using mixture model cutoffs.
```{r}
barcodes_to_remove <- map(groups, function(i) {
  merged[[i]]@meta.data %>% 
    rownames_to_column() %>% 
    filter(seurat_clusters %in% low_qc_clusters[[i]]) %>% 
    mutate(cell_barcode = str_sub(rowname, end = -3)) %>% 
    select(cell_barcode, orig.ident)
}) %>% bind_rows()

barcodes_to_remove
```

### QC filter using k-means groups and additional cutoffs based on mixture models
Set up mixture model cutoffs

Mixture model on QC metrics
```{r fig.height=8, fig.width=16}
map(seurats, function(seurat) {
  
  data_input <- seurat@meta.data$nFeature_RNA %>% log10()
  mixture <- mixture_model(data_input)
  
  mixture$plot +
    ggtitle(seurat@project.name) 
  
  
}) %>% wrap_plots(nrow = 2)

```
Ribo fraction
```{r fig.height=8, fig.width=16}
map(seurats, function(seurat) {
  data_input <- seurat@meta.data$percent_ribo %>% log10()
  
  mixture <- mixture_model(data_input)
  mixmod <- mixture$model
  
  mixture$plot +
    ggtitle(seurat@project.name) 
  
  
}) %>% wrap_plots(nrow = 2)
```

Mito fraction
```{r fig.height=8, fig.width=16}
map(seurats, function(seurat) {
  data_input <- seurat@meta.data$percent_mito 
  
  mixture <- mixture_model(data_input)
  
  mixture$plot +
    ggtitle(seurat@project.name) 
  
  
}) %>% wrap_plots(nrow = 2)
```

## Visualize
```{r fig.height=16, fig.width=14}
map(seurats, function(seurat) {
  barcodes_to_remove_subset <- barcodes_to_remove %>% 
    filter(orig.ident == seurat@project.name) %>% 
    pull(cell_barcode)
  
  # nFeature
  feature_mixture <- mixture_model(seurat@meta.data$nFeature_RNA %>% log10())
  
  # ribo
  ribo_mixture <- mixture_model(seurat@meta.data$percent_ribo %>% log10())
  
  # mito
  mito_mixture <- mixture_model(seurat@meta.data$percent_mito)
  
  barcodes_to_keep <- rownames_to_column(seurat@meta.data, 'cell_barcode') %>% 
    filter(
      !(cell_barcode %in% barcodes_to_remove_subset)) %>% 
    pull(cell_barcode)
  
  seurat@meta.data %>% 
    rownames_to_column('cell_barcode') %>% 
    mutate('QC' = ifelse(cell_barcode %in% barcodes_to_keep, 'Include', 'Exclude')) %>% 
    ggplot(aes(x = percent_ribo,
               y = log10(nFeature_RNA),
               color = QC)) + 
    geom_point(alpha = 0.8, size = 0.25) +
    scale_color_few() +
    ggtitle(seurat@project.name) + 
    theme(legend.position = 'bottom') +
    facet_wrap(~QC, nrow = 2)
  
}) %>% wrap_plots(nrow = 2)
```

Filter using QC as well as mixture model cutoffs
```{r}
# seurats <- seurats %>% map(function(seurat) {
#   
#    barcodes_to_remove_subset <- barcodes_to_remove %>% 
#      filter(orig.ident == seurat@project.name) %>% 
#      pull(cell_barcode)
#    
#    # nFeature
#    feature_mixture <- mixture_model(seurat@meta.data$nFeature_RNA %>% log10())
#   
#    # ribo
#    ribo_mixture <- mixture_model(seurat@meta.data$percent_ribo %>% log10())
#    
#    # mito
#    mito_mixture <- mixture_model(seurat@meta.data$percent_mito)
#    
#    barcodes_to_keep <- rownames_to_column(seurat@meta.data, 'cell_barcode') %>% 
#      filter(
#        qc_cluster != 'QC3',
#        !(cell_barcode %in% barcodes_to_remove_subset)) %>% 
#      pull(cell_barcode)
#    
#    subset(seurat, cells = barcodes_to_keep)
# })
# 
# seurats
```
### Final visualization
```{r fig.height=8, fig.width=16}
map(seurats, function(seurat) {
  seurat@meta.data %>% 
  ggplot(aes(x = percent_ribo,
             y = log10(nFeature_RNA),
             color = qc_cluster)) + 
  geom_point(alpha = 0.5, size = 0.25) +
  scale_color_few() +
    ggtitle(seurat@project.name) + 
    theme(legend.position = 'bottom')
}) %>% wrap_plots(nrow = 2)

```
## Export
```{r}
seurats %>% write_rds(file.path(data_dir, 'unified_cellranger_qc.rds'))
```


## Session info
```{r}
sessionInfo()
```

