---
title: "Integrate Perturb-Seq data"
author: "David Wu"
output: html_notebook
---
## Purpose
Use CCA to integrate batches

## Setup 
Set working directory to project directory
```{r setup}
require(knitr)
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```
### Optionally convert this Rmd to R script 
```{r eval=FALSE}
current_file <- rstudioapi::getActiveDocumentContext()$path
output_file <- stringr::str_replace(current_file, '.Rmd', '.R')
knitr::purl(current_file, output = output_file)
file.edit(output_file)
```

### Load custom themes and functions
```{r, message=FALSE}
source('analysis/universal/themes.R') # themes for all project components

source('analysis/universal/functions.R') # themes for all project components
theme_set(theme_publication())

source('analysis/perturbseq/scripts/functions.R')  # functions for this component; may be empty
```

### Directories
```{r, message=FALSE}
analysis_dir <- 'analysis/perturbseq/output/008_integration' # analysis file output directory
data_dir <- 'data/perturbseq/derived/seurat' # data file output directory

dir.create(analysis_dir, showWarnings = FALSE)
dir.create(data_dir, showWarnings = FALSE)
```

### Load libraries
```{r, message=FALSE}
library(tidyverse)
library(Seurat)
```

### Import data
Given results of QC analysis, perform downstream pipeline using all filtered singlets (but keep QC metrics)
```{r}
seurats <- read_rds('data/perturbseq/derived/seurat/unified_cellranger_qc.rds')
seurats
```


## Integration
### Set up
```{r}
dims_to_integrate <- 1:30
```

### Integrate
Use reference-based since all samples are similar
Use largest sample as reference (for each condition)
Much faster to run on command line
SCTransform in conjunction with CCA
```{r}
tic()

seurat_list <- seurats %>% 
  map(SCTransform, verbose = TRUE)

largest_sample <- map(seurat_list, ncol) %>% unlist() %>% sort(decreasing = TRUE) %>% names() %>% .[1]
reference_sample <- which(names(seurat_list) == largest_sample)

integration_features <- SelectIntegrationFeatures(object.list = seurat_list,
                                                  nfeatures = 5000)

seurat_list <- PrepSCTIntegration(object.list = seurat_list, 
                                  assay = 'SCT',
                                  anchor.features = integration_features,
                                  verbose = TRUE)

anchors <- FindIntegrationAnchors(object.list = seurat_list, # this step takes a while
                                  normalization.method = 'SCT',
                                  anchor.features = integration_features,
                                  dims = dims_to_integrate,
                                  reference = reference_sample,
                                  verbose = TRUE)

integrated <- IntegrateData(anchorset = anchors,
                            normalization.method = 'SCT',
                            dims = dims_to_integrate,
                            verbose = TRUE)

integrated


toc()

integrated <- integrated %>% CellCycleScoring(s.features = cc.genes$s.genes, g2m.features = cc.genes$g2m.genes)

# Export
integrated %>% write_rds(file.path(data_dir, 'unified_cellranger_sct.rds'))


Sys.time()
```


## Session info                     
```{r}
sessionInfo()
```

