---
title: "Clustering of Perturb-Seq data"
author: "David Wu"
output: html_notebook
---
## Purpose
Run clustering on PCA data

## Setup 
Set working directory to project directory
```{r setup}
require(knitr)
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

### Optionally convert this Rmd to R script 
```{r eval=FALSE}
current_file <- rstudioapi::getActiveDocumentContext()$path
output_file <- stringr::str_replace(current_file, '.Rmd', '.R')
knitr::purl(current_file, output = output_file)
file.edit(output_file)
```

### Load custom themes and functions
```{r, message=FALSE}
source('analysis/universal/themes.R') # themes for all project components

source('analysis/universal/functions.R') # themes for all project components
theme_set(theme_publication())

source('analysis/perturbseq/scripts/functions.R')  # functions for this component; may be empty
```

### Directories
```{r, message=FALSE}
analysis_dir <- 'analysis/perturbseq/output/010_clustering' # analysis file output directory
data_dir <- 'data/perturbseq/derived/seurat' # data file output directory

dir.create(analysis_dir, showWarnings = FALSE)
dir.create(data_dir, showWarnings = FALSE)
```

### Load libraries
```{r, message=FALSE}
library(tidyverse)
library(tictoc)
library(Seurat)
library(leiden)
```

## Globals
Quick umap parameters for test plotting; see 011_dimensionality_reduction for final UMAP parameters
```{r}
calculate_dims <- 50
significant_dims <- 1:30
k <- 30
neighbors <- 25
min_dist <- 0.05
n_epochs <- 250
umap_seed <- 737697
algorithm <- 'both'
louvain_resolution <- c(0.6, 0.8, 1.0)
leiden_resolution <- c(0.6, 0.8, 1.0)
```

# Clustering on SCT data
## Import integrated object
```{r}
merged <- read_rds('data/perturbseq/derived/seurat/unified_cellranger_sct.rds')
merged
```
### Scale data
```{r}
# check if SCT
if(any(str_detect(names(merged@assays), 'SCT'))) {
  scale_data <- FALSE
  assay_name <- 'SCT'
} else {
  scale_data <- TRUE
  assay_name <- 'RNA'
}

if(scale_data) {
  merged <- merged %>% 
    ScaleData()
  
  merged <- merged %>% ScaleData(assay = 'integrated')
}

merged
```


### PCA
```{r}
merged <- merged %>% RunPCA(assay = 'integrated',
                             npcs = calculate_dims)

merged %>% ElbowPlot(ndims = calculate_dims)
```

### Clustering 
```{r}
tic()
merged <- merged %>% FindNeighbors(assay = 'integrated',
                                   dims = significant_dims, 
                                   k.param = neighbors)

toc()

# louvain alg
if(algorithm == 'louvain' | algorithm == 'both') {
  tic()
  merged <- merged %>% FindClusters(assay = 'integrated', algorithm = 1, resolution = louvain_resolution)
  
  names(merged@meta.data)[which(names(merged@meta.data) == paste0('integrated_snn_res.', max(louvain_resolution)))] <- 'louvain_cluster_hi'
  names(merged@meta.data)[which(names(merged@meta.data) == paste0('integrated_snn_res.', median(louvain_resolution)))] <- 'louvain_cluster_mid'
  names(merged@meta.data)[which(names(merged@meta.data) == paste0('integrated_snn_res.', min(louvain_resolution)))] <- 'louvain_cluster_lo'

toc()  
} 

# leiden alg; need to assign manually
if (algorithm == 'leiden' | algorithm == 'both') {
  
  tic()
  merged <- merged %>% FindClusters(assay = 'integrated', algorithm = 4, resolution = leiden_resolution, method = 'igraph')
  names(merged@meta.data)[which(names(merged@meta.data) == paste0('integrated_snn_res.', max(leiden_resolution)))] <- 'leiden_cluster_hi'
  names(merged@meta.data)[which(names(merged@meta.data) == paste0('integrated_snn_res.', median(leiden_resolution)))] <- 'leiden_cluster_mid'
  names(merged@meta.data)[which(names(merged@meta.data) == paste0('integrated_snn_res.', min(leiden_resolution)))] <- 'leiden_cluster_lo'
  
  toc()
}
```



### Visualization
Test UMAP for a quick check on clustering; optimize parameters in dimensionality reduction notebook
```{r fig.height=6, fig.width=6}
tic()
test <- merged %>% RunUMAP(assay = 'integrated',
                           dims = significant_dims, 
                           reduction = 'pca', 
                           reduction.name = 'umap',
                           n.components = 2, 
                           n.neighbors = k, 
                           min.dist = min_dist,
                           n.epochs = n_epochs,
                           seed.use = umap_seed)
toc()

test %>% seurat_plot(reduction = 'umap', feature = 'ident', do.label = T) %>% wrap_plots()
```


## Visualize Cell cycle
```{r}
test %>% seurat_features(reduction = 'umap', gene.list = 'S.Score') 
```
```{r}
test %>% seurat_features(reduction = 'umap', gene.list = 'G2M.Score') 
```

## Export
```{r}
merged %>% write_rds(file.path(data_dir, 'unified_cellranger_sct_clustered.rds'))
```


## Session info
```{r}
sessionInfo()
```

