---
title: "Dimensionality Reduction of Perturb-Seq data"
author: "David Wu"
output: html_notebook
---
## Purpose
Run dimensionality reduction 

Set working directory to project directory
```{r setup}
require(knitr)
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

### Optionally convert this Rmd to R script 
```{r eval=FALSE}
current_file <- rstudioapi::getActiveDocumentContext()$path
output_file <- stringr::str_replace(current_file, '.Rmd', '.R')
knitr::purl(current_file, output = output_file)
file.edit(output_file)
```

### Load custom themes and functions
```{r, message=FALSE}
source('analysis/universal/themes.R') # themes for all project components

source('analysis/universal/functions.R') # themes for all project components
theme_set(theme_publication())

source('analysis/perturbseq/scripts/functions.R')  # functions for this component; may be empty
```

### Directories
```{r, message=FALSE}
analysis_dir <- 'analysis/perturbseq/output/011_dimensionality_reduction' # analysis file output directory
data_dir <- 'data/perturbseq/derived/seurat' # data file output directory

dir.create(analysis_dir, showWarnings = FALSE)
dir.create(data_dir, showWarnings = FALSE)
```

### Load libraries
```{r, message=FALSE}
library(tidyverse)
library(Seurat)
library(plotly)
```

### Import feature table
```{r}
feature_reference <- read_csv('data/perturbseq/derived/pipestance/features.csv')
feature_reference %>% head()
```
### Import screen data
```{r}
master_minimal <- read_tsv('analysis/integration/output/00_integrate/master_minimal.tsv.gz')
master_minimal %>% head()
```

# SCT
## Import integrated object
```{r}
tic()
integrated <- read_rds('data/perturbseq/derived/seurat/unified_cellranger_sct_clustered.rds')
toc()
integrated
``` 

## Globals
```{r}
significant_dims <- 1:30
k <- 30
min_dist <- 0.1
n_epochs <- 250
umap_seed <- 20200522

tsne_seed <- umap_seed
theta <- 0.2
iter <- 1000

which_assay <- 'integrated'
```

```{r}
tic()
integrated <- integrated %>% 
  RunPCA(npcs = max(significant_dims) + 10,
         assay = which_assay, 
         verbose = FALSE)
toc()
```

### UMAP
```{r fig.height=6, fig.width=6}
tic()
reduced <- integrated %>% 
  RunUMAP(
          dims = significant_dims, 
          assay = which_assay,
          reduction = 'pca', 
          reduction.name = 'umap',
          n.components = 2, 
          n.neighbors = k, 
          min.dist = min_dist,
          n.epochs = n_epochs,
          seed.use = umap_seed)

toc()

all_umap <- reduced %>% seurat_plot(reduction = 'umap', feature = 'ident', do.label = T, size = 0.05, alpha = 0.2) %>% wrap_plots()

ggsave(plot = all_umap, 
       path = analysis_dir, 
       filename = 'unified_cellranger_sct_umap.png', 
       h = 3, w = 3)

all_umap
```


### 3D UMAP
```{r eval=TRUE}
reduced <- reduced %>% 
  RunUMAP(
          assay = which_assay,
          dims = significant_dims, 
          reduction = 'pca', 
          reduction.name = 'umap3d',
          n.components = 3, 
          n.neighbors = k, 
          min.dist = min_dist,
          n.epochs = n_epochs,
          seed.use = umap_seed)

```


```{r eval=FALSE}
seurat_3d(reduced)
```


### t-SNE
```{r fig.height=12, fig.width=12, eval=FALSE}
tic()
reduced <- reduced %>% 
  RunTSNE(
          dims = significant_dims, 
          assay = which_assay,
          reduction = 'pca', 
          tsne.method = 'FIt-SNE', 
          fast_tsne_path = '/home/software/FIt-SNE/bin/fast_tsne', 
          max_iter = iter, 
          nthreads = 24,
          theta = theta, 
          perplexity = k,
          seed.use = tsne_seed)
  })
toc()
```

### Low resolution clustering to filter for multiple trajectory inference
Perform low resolution clustering on UMAP coordinates for downstream pseudotime analysis
```{r}
trajectory_clustering <- reduced %>% 
  FindNeighbors(reduction = 'umap',
                dims = 1:2, 
                k = 100)

trajectory_clustering <- trajectory_clustering %>% 
  FindClusters(resolution = 0.3)

seurat_plot(trajectory_clustering, feature = 'seurat_clusters')
```


```{r}
trajectory <- case_when(
  trajectory_clustering$seurat_clusters %in% c(9, 3, 6, 5) ~ 'Cell Cycle',
  trajectory_clustering$seurat_clusters %in% c(10, 1, 0, 2, 16, 4) ~ 'NSC',
  trajectory_clustering$seurat_clusters %in% c(11, 7, 12) ~ 'Non-NSC',
  TRUE ~ 'Other')

reduced$low_res_cluster <- trajectory_clustering$seurat_clusters
reduced$trajectory <- trajectory

seurat_plot(reduced, feature = 'trajectory')
```

## Reset to RNA assay
```{r}
DefaultAssay(integrated) <- 'RNA'
integrated <- integrated %>% NormalizeData()
    
```

## Add final metadata
```{r}
metadata_table <- integrated@meta.data %>% 
  rownames_to_column('barcode') %>%
  left_join(master_minimal %>% 
              dplyr::select(guide_target = feature_id,
                            library,
                            primary_pheno,
                            diff_pheno,
                            diff_hit_direction) 
            %>% unique()) %>% 
  as_tibble()

metadata_table
```

```{r}
data_to_add <- metadata_table %>% 
  dplyr::select(barcode,
                library,
                primary_pheno,
                diff_pheno,
                diff_hit_direction) %>% 
  mutate_if(is.numeric, ~replace(., is.na(.), 0)) %>% 
  mutate_if(is.character, ~replace(., is.na(.), 'Control')) %>% 
  column_to_rownames('barcode') 

reduced <- AddMetaData(reduced, data_to_add)

```

## Export
```{r}
reduced %>% write_rds(file.path(data_dir, 'unified_cellranger_sct_reduced.rds'))
```

## Session info
```{r}
sessionInfo()
```

