---
title: "Perturbation diversity"
author: "David Wu"
output: html_notebook
---
## Purpose
Analyze diverse perturbation effects

Set working directory to project directory
```{r setup}
require(knitr)
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

### Optionally convert this Rmd to R script 
```{r eval=FALSE}
current_file <- rstudioapi::getActiveDocumentContext()$path
output_file <- stringr::str_replace(current_file, '.Rmd', '.R')
knitr::purl(current_file, output = output_file)
file.edit(output_file)
```
### Load custom themes and functions
```{r, message=FALSE}
source('analysis/universal/themes.R') # themes for all project components

source('analysis/universal/functions.R') # themes for all project components
theme_set(theme_publication())

source('analysis/perturbseq/scripts/functions.R')  # functions for this component; may be empty
```

### Output directory
```{r}
analysis_dir <- 'analysis/perturbseq/output/cell_state_frequency'
figures_dir <- 'figures/perturbseq'

dir.create(analysis_dir, showWarnings = FALSE)
dir.create(figures_dir, showWarnings = FALSE)
```

```{r}
library(extrafont)
```

### Load files
## Coordinates and metadata
```{r}
DIR <- 'analysis/perturbseq/output/data_export/'
pca_file <- 'pca_coordinates.tsv.gz'
umap_file <- 'umap_coordinates.tsv.gz'
meta_file <- 'coordinate_metadata.tsv.gz'

dr <- list('pca' = read_tsv(file.path(DIR, pca_file)),
           'umap' = read_tsv(file.path(DIR, umap_file)))
           
meta <- read_tsv(file.path(DIR, meta_file))

coordinates_hclust <- read_tsv('analysis/perturbseq/output/density_clustering/coordinates_hclust.tsv.gz')
```
```{r}
joined_data <- coordinates_hclust %>% 
  left_join(meta %>% dplyr::select(target = guide_target, library, diff_hit_direction, diff_full_status, diff_pheno, locus_ambiguity) %>% unique()) %>% 
  filter(library == 'Coding' | locus_ambiguity == 'Non-ambiguous', diff_full_status != 'Neighbor hit', diff_hit_direction != 'Non-hit') 

joined_data
```

### Fully annotate table
```{r}
nsc_states <- c(16, 12, 9, 23, 24, 3, 14, 27, 20, 21, 13, 6, 28)
cycling_states <- c(18, 4, 10, 25, 15, 8, 5)
ectoderm_states <- c(2, 11, 19, 1)
other_states <- c(26, 17, 7, 29, 22)

nonneural_states <- c(ectoderm_states, other_states)
nonneural_states <- ectoderm_states
apoptotic_states <- c(20)

state_types <- joined_data %>% 
  group_by(target, library, diff_hit_direction, hc_cluster) %>% 
  mutate(max_z = max(z)) %>% 
  group_by(target, library, diff_hit_direction, hc_cluster, max_z) %>% 
  summarize(density = quantile(z, probs = 0.5)) %>% 
  unique() %>% 
  mutate(trajectory = case_when(
    hc_cluster %in% nsc_states ~ 'NSC',
    hc_cluster %in% cycling_states ~ 'Cell Cycle',
    hc_cluster %in% ectoderm_states ~ 'Ectoderm',
    hc_cluster %in% other_states ~ 'Other'),
    neural = ifelse(hc_cluster %in% nonneural_states, 'Non-neural', 'Neural'),
    apoptosis = ifelse(hc_cluster %in% apoptotic_states, 'Apoptosis', 'Non-apoptosis')) %>% 
  group_by(target) %>% 
  add_count(name = 'n_states') %>% 
  mutate(states = ifelse(
    n_states > 2, 
    '3+',
    '1-2'),
    density_zscore = scale(density)) %>% 
  ungroup() 

state_types  
```

### For negative hits, what are the most frequent patterns of preventing neural induction?
```{r}
frequency_table <- state_types %>%   
  filter(diff_hit_direction == 'Negative Hit') %>% 
  group_by(library, hc_cluster) %>% 
  add_count(name = 'cell_state_frequency') %>% 
  arrange(-cell_state_frequency) %>% 
  ungroup()

frequency_table
  
```

```{r}
frequency_table %>% 
  ggplot(aes(x = reorder(hc_cluster, -cell_state_frequency),
             y = cell_state_frequency,
             fill = library)) +
  geom_bar(stat = 'identity',
           position = 'dodge') +
  facet_wrap(~library)
```
```{r}
frequency_table %>% 
  group_by(library, hc_cluster) %>% 
  tally() %>% 
  arrange(-n)
  
```
```{r}
top_states <- frequency_table %>% 
  group_by(target) %>% 
  top_n(1, density) %>% 
  ungroup() %>% 
  group_by(library) %>% 
  add_count(hc_cluster, name = 'top_state_frequency') %>% 
  arrange(-top_state_frequency)


intermediate <- top_states %>% filter(hc_cluster %in% c(2, 6, 13)) %>% pull(target) %>% unique()
apoptotic <- top_states %>% filter(hc_cluster %in% c(20, 27)) %>% pull(target) %>% unique()
nonneural <- top_states %>% filter(hc_cluster %in% ectoderm_states) %>% pull(target) %>% unique()
cycling <- top_states %>% filter(hc_cluster %in% cycling_states) %>% pull(target) %>% unique()
nsc <-  top_states %>% filter(hc_cluster %in% nsc_states) %>% pull(target) %>% unique()

category_tally <- 
  top_states %>% 
  select(target, library, n_states, states, top_state = hc_cluster, top_state_frequency) %>%
  mutate(category = case_when(target %in% intermediate ~ 'Intermediate',
                              target %in% apoptotic  ~ 'Apoptosis',
                              target %in% cycling ~ 'Cycling', 
                              target %in% nonneural ~ 'Non-CNS',
                              target %in% nsc ~ 'NSC',
                              TRUE ~ 'Other'),
         category = factor(category))
category_tally
```
```{r}
library_totals <- category_tally %>% select(target, library) %>% unique() %>% group_by(library) %>% tally(name = 'total')
library_totals
```


```{r}
category_summary <- category_tally %>% 
  select(library, category) %>% 
  table() %>% 
  as_tibble() %>% 
  left_join(library_totals) %>% 
  mutate(frac = round(n / total, 3))

category_summary
```
```{r}
category_summary %>% 
  filter(category %in% c('Intermediate', 'Apoptosis', 'Non-CNS')) %>% 
  ggplot(aes(x = category,
             y = frac,
             fill = library)) +
  geom_bar(stat = 'identity',
           position = 'dodge') +
  scale_fill_discrete(drop=FALSE) +
          scale_x_discrete(drop=FALSE) +
  labs(x = '',
       y = 'Proportion') +
  theme(legend.position = 'top')
```

### Abortive vs permissive state
```{r}
abortive <- frequency_table %>% 
  group_by(target) %>% 
  top_n(1, density) %>% 
  filter(hc_cluster %in% c(21, 14, 3, 13, 6, 20))

abortive %>% 
  ungroup() %>% 
  filter(library == 'Coding') %>% 
  add_count(hc_cluster) %>% 
  arrange(-n)
```
```{r}
frequency_table %>% 
  group_by(target) %>% 
  top_n(1, density) %>% 
  filter(library == 'Coding') %>% 
  ungroup() %>% 
  add_count(hc_cluster) %>% 
  arrange(-n) %>% 
  select(target, hc_cluster, n_states, n)
```


```{r}
abortive %>% group_by(library) %>% tally() %>% 
  left_join(library_totals) %>% 
  mutate(frac = n/total)
```

```{r}
permissive <- 
  frequency_table %>% 
  filter(states == '3+') %>% 
  group_by(target, trajectory) %>% 
  add_count(name = 'n_trajectories') %>% 
  ungroup() %>% 
  select(target, library, n_states, n_trajectories) %>% 
  unique() %>% 
  arrange(n_trajectories)

permissive %>% 
  filter(n_trajectories > 1) %>% 
  group_by(library) %>% 
  tally() %>% 
  left_join(library_totals) %>% 
  mutate(frac = n/total)
```

### Trajectories per target
```{r}
n_trajectories <- frequency_table %>% 
  filter(trajectory != 'Other') %>% 
  select(target, trajectory) %>% 
  unique() %>% 
  group_by(target) %>% 
  add_count(name = 'n_trajectories') %>% 
  select(-trajectory) %>% 
  ungroup() %>% 
  unique()

n_trajectories
```

```{r}
plot_input <- frequency_table %>% 
  left_join(n_trajectories) %>% 
  mutate(n_trajectories = ifelse(is.na(n_trajectories), 1, n_trajectories)) %>% # for hits in "Other" trajectory
  group_by(target) %>% 
  top_n(1, density) %>% 
  mutate(abortive = ifelse(hc_cluster %in% c(21, 14, 3, 13, 6, 20, 27), 'Yes', 'No')) %>% 
  select(target, library, top_state = hc_cluster, n_states, states, n_trajectories, abortive) %>% 
  mutate(group = case_when(
    top_state %in% c(21, 14, 3, 13, 6) ~ 'Stalled',
    top_state %in% c(20, 27) ~ 'Apoptosis',
    n_trajectories > 1 ~ 'Multiple trajectories',
    TRUE ~ 'Other'
  ))
  # mutate(group = case_when(
  #   n_trajectories == 1 & top_state %in% c(21, 14, 3, 13, 6) ~ 'Stalled',
  #   n_trajectories == 1 & top_state %in% c(20, 27) ~ 'Apoptosis',
  #   n_trajectories > 1 ~ 'Multiple trajectories',
  #   TRUE ~ 'Other'
  # ))

plot_input
```

```{r}
totals <- plot_input %>% 
  group_by(library) %>% 
  tally(name ='total')

group_tally <- plot_input %>% 
  group_by(group, library) %>% 
  tally() %>% 
  ungroup() %>% 
  left_join(totals) %>% 
  mutate(frac = round(n/total, 3))
 
group_tally
```

### Test plots
```{r}
p_pie <- group_tally %>% 
  group_by(library) %>% 
  mutate(group = factor(group, levels = c('Stalled', 'Apoptosis', 'Other', 'Multiple trajectories'))) %>% 
  arrange(group) %>% 
  mutate(prop = frac) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop ) %>% 
  mutate(label = paste0(round(prop * 100), '%')) %>% 
  ggplot(aes(x = '',
             y = frac,
             fill = group)) +
  geom_bar(stat = 'identity', width = 1, color = 'black', alpha = 1) +
  geom_text(aes(y = (1 - ypos), label = label), color = 'black') + 
  coord_polar('y', start = 0) +
  facet_wrap(~library) +
  #scale_fill_few() + 
  scale_fill_manual(values = c('dodgerblue3', 'dodgerblue4','grey90', 'orangered')) +
  theme_void() 

save_figure(p_pie,
            dir = 'figures/perturbseq',
            filename = 'pie_abortive_vs_permissive')

p_pie
```


### Enrichment stats
```{r}
enrich_input <- plot_input %>% select(target, library, group) %>% 
  filter(group != 'Other')
  
  
in_category <- enrich_input %>% filter(group %in% c('Stalled', 'Apoptosis')) %>% pull(target)
in_category <- enrich_input %>% filter(group %in% 'Multiple trajectories') %>% pull(target)
is_coding <- enrich_input %>% filter(library == 'Coding') %>% pull(target)

enrich_res <- enrichment_test(list1 = is_coding,
                              list2 = in_category,
                              cat1 = c('Coding', 'lncRNA'),
                              cat2 = c('In', 'Out'),
                              background = enrich_input$target,
                              print = T)$fisher
```


### Separate calculations
```{r}
abortive_tally <- frequency_table %>% 
  group_by(target, trajectory) %>% 
  add_count(name = 'n_trajectories') %>% 
  group_by(target) %>% 
  top_n(1, density) %>% 
  mutate(abortive = ifelse(hc_cluster %in% c(21, 14, 3, 13, 6, 20, 27), 'Yes', 'No'),
         permissive = ifelse(n_trajectories > 1, 'Yes', 'No')) %>% 
  select(target, library, top_state = hc_cluster, n_states, states, n_trajectories, abortive, permissive) %>% 
  mutate(group = case_when(
    top_state %in% c(21, 14, 3, 13, 6) ~ 'Stalled',
    top_state %in% c(20, 27) ~ 'Apoptosis',
    TRUE ~ 'Other'
  ))

abortive_tally
```


```{r}
abortive_totals <- abortive_tally %>% 
  group_by(library) %>% 
  tally(name = 'total')

abortive_totals

abortive_prop <- abortive_tally %>% 
  group_by(library, abortive) %>% 
  tally() %>% 
  ungroup() %>% 
  left_join(abortive_totals) %>% 
  mutate(frac = round(n/total, 3))

abortive_prop
```
```{r}
permissive_prop <- abortive_tally %>% 
  group_by(library, permissive) %>% 
  tally() %>% 
  ungroup() %>% 
  left_join(abortive_totals) %>% 
  mutate(frac = round(n/total, 3))
permissive_prop
```


### Pie charts separate
```{r}
p_pie1 <- abortive_prop %>% 
  group_by(library) %>% 
  mutate(group = factor(abortive, levels = c('Yes', 'No'))) %>% 
  arrange(group) %>% 
  mutate(prop = frac) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop ) %>% 
  mutate(label = paste0(round(prop * 100), '%')) %>% 
  ggplot(aes(x = '',
             y = frac,
             fill = group)) +
  geom_bar(stat = 'identity', width = 1, color = 'black', alpha = 1) +
  geom_text(aes(y = (1 - ypos), label = label), color = 'black') + 
  coord_polar('y', start = 0) +
  facet_wrap(~library) +
  #scale_fill_few() + 
  scale_fill_manual(values = c('dodgerblue3', 'grey90')) +
  theme_void() + 
  theme(legend.position = 'none',
        plot.title = element_text(hjust = 0.5)) +
  labs(title = 'Abortive')

p_pie2 <- permissive_prop %>% 
  group_by(library) %>% 
  mutate(group = factor(permissive, levels = c('Yes', 'No'))) %>% 
  arrange(group) %>% 
  mutate(prop = frac) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop ) %>% 
  mutate(label = paste0(round(prop * 100), '%')) %>% 
  ggplot(aes(x = '',
             y = frac,
             fill = group)) +
  geom_bar(stat = 'identity', width = 1, color = 'black', alpha = 1) +
  geom_text(aes(y = (1 - ypos), label = label), color = 'black') + 
  coord_polar('y', start = 0) +
  facet_wrap(~library) +
  #scale_fill_few() + 
  scale_fill_manual(values = c('orangered', 'grey90')) +
  theme_void() +
  theme(legend.position = 'none',
        plot.title = element_text(hjust = 0.5)) +
  labs(title = 'Permissive')

p_pies <- p_pie1 + p_pie2

save_figure(p_pies,
            dir = 'figures/perturbseq',
            filename = 'pie_separate_abortive_vs_permissive')

p_pies
```
### Rho vs multi state
```{r}
rho_analysis <- plot_input %>% 
  left_join(joined_data %>% 
              select(target, diff_hit_direction, diff_full_status, diff_pheno) %>% 
              unique())

rho_analysis
```
```{r}
cor_input <- rho_analysis %>% filter(library == 'Coding')
cor.test(cor_input$n_trajectories,
         cor_input$diff_pheno,
         method = 'pearson')
```
```{r}
cor_input <- rho_analysis %>% filter(library == 'lncRNA')
cor.test(cor_input$n_trajectories,
         cor_input$diff_pheno,
         method = 'pearson')
```

```{r}
cor_input <- rho_analysis
cor.test(cor_input$n_trajectories,
         cor_input$diff_pheno,
         method = 'pearson')
```
```{r}
p_rho <- rho_analysis %>% 
  ggplot(aes(x = n_trajectories,
             y = diff_pheno)) +
  geom_point(alpha = 1) +
  geom_smooth(method = 'lm', se = FALSE) +
  labs(x = '# trajectories',
       y = 'rho') + 
  ylim(c(-3.1, 0))

save_figure(p_rho,
            dir = 'figures/perturbseq',
            filename = 'cor_rho_vs_trajectories')


p_rho
```
```{r}
p_rho_sina <- rho_analysis %>% 
  filter(group %in% c('Stalled', 'Apoptosis', 'Multiple trajectories')) %>% 
  ggplot(aes(x = abortive,
             y = diff_pheno,
             group = abortive)) +
  geom_sina(alpha = 0.5) +
  #geom_boxplot() +
  geom_errorbar(stat = "summary",
                width=0.2,
                color = 'orangered') +
  ylim(c(-3.1, 0)) +
  labs(x = '',
       y = 'Rho')


save_figure(p_rho_sina,
            dir = 'figures/perturbseq',
            filename = 'sina_rho')

p_rho_sina
```
```{r}
test_input <- rho_analysis %>% 
  filter(group %in% c('Stalled', 'Apoptosis', 'Multiple trajectories')) 

t.test(rho_analysis %>% filter(group %in% c('Stalled', 'Apoptosis')) %>% pull(diff_pheno),
            rho_analysis %>% filter(group %in% 'Multiple trajectories') %>% pull(diff_pheno))
```

```{r}
rho_analysis %>% 
  ggplot(aes(x = n_trajectories,
             y = diff_pheno)) +
  geom_point(alpha = 1) +
  geom_smooth(method = 'lm', se = FALSE) +
  labs(x = '# trajectories',
       y = 'rho') + 
  facet_wrap(~library,
             nrow = 1) +
  ylim(c(-3.1, 0))
```



## Session info
```{r}
sessionInfo()
```
