---
title: "Distribution statistics"
author: "David Wu"
output: html_notebook
---
## Purpose
Compare distributions of sgRNA categories using multivariate KS test

## Setup 
Set working directory to project directory
```{r setup}
require(knitr)
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

### Optionally convert this Rmd to R script 
```{r eval=FALSE}
current_file <- rstudioapi::getActiveDocumentContext()$path
output_file <- stringr::str_replace(current_file, '.Rmd', '.R')
knitr::purl(current_file, output = output_file)
file.edit(output_file)
```

### Load custom themes and functions
```{r, message=FALSE}
source('analysis/universal/themes.R') # themes for all project components

source('analysis/universal/functions.R') # themes for all project components
theme_set(theme_publication())

source('analysis/perturbseq/scripts/functions.R')  # functions for this component; may be empty
```

### Output directory
```{r}
analysis_dir <- 'analysis/perturbseq/output/distribution_statistics'
figures_dir <- 'figures/perturbseq'

dir.create(analysis_dir, showWarnings = FALSE)
dir.create(figures_dir, showWarnings = FALSE)
```

### Custom functions for this notebook

### Load libraries
```{r, message=FALSE}
library(tidyverse)
library(GSAR)
library(tictoc)
library(parallel)
library(Peacock.test)
```

# Load directories

```{r}
## -------------------------------------------------------------------------------------------------------
DIR <- 'analysis/perturbseq/output/data_export/'
pca_file <- 'pca_coordinates.tsv.gz'
umap_file <- 'umap_coordinates.tsv.gz'
meta_file <- 'coordinate_metadata.tsv.gz'

dr <- list('pca' = read_tsv(file.path(DIR, pca_file)),
           'umap' = read_tsv(file.path(DIR, umap_file)))
           
meta <- read_tsv(file.path(DIR, meta_file))

combined <- dr$umap %>% 
  left_join(meta) %>% 
  filter(diff_full_status != 'Neighbor hit',
         locus_ambiguity == 'Non-ambiguous' | library == 'Coding') %>% 
  select(UMAP_1, UMAP_2, library, primary_pheno, diff_hit_direction)
combined %>% head()
```

```{r}
rm(dr)
rm(meta)
gc()
```


## Statistics to compare distributions
### Multivariate KS test
Run in parallel as each computation can take several hours
```{r}
groups <- c('non', 'pos', 'neg')

tic()
mclapply(groups,
                   mc.cleanup = TRUE,
                   mc.cores = length(groups),
                   function(i) {

           if(i == 'non') {
             test_subset <- combined %>% filter(primary_pheno %in% c('None', 'Control'))
           } else if (i == 'pos') {
             test_subset <- combined %>% filter(diff_hit_direction %in% c('Positive Hit', 'Control'))
           }  else if (i == 'neg') {
             test_subset <- combined %>% filter(diff_hit_direction %in% c('Negative Hit', 'Control'))
           } 
           
           test_subset <- test_subset %>% 
             mutate(group = ifelse(primary_pheno == 'Control', 1, 2)) %>% 
             group_by(group) %>% 
             sample_frac(0.5) %>% 
             ungroup() %>% 
             select(UMAP_1, UMAP_2, group)
             
           tic()
           ks <- KStest(object = t(as.matrix(test_subset[,c('UMAP_1', 'UMAP_2')])),
                            group = test_subset$group,
                            nperm = 1000,
                            pvalue.only = FALSE)
           toc()
           
           ks %>% write_rds(file.path(analysis_dir, paste0('ks_', i, '.rds')))
           
           ks
})

toc()
Sys.time()


```

