---
title: "Heatmaps"
author: "David Wu"
output: html_notebook
---
## Purpose
Heatmaps of perturbations and gene expression

Set working directory to project directory
```{r setup}
require(knitr)
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

### Optionally convert this Rmd to R script 
```{r eval=FALSE}
current_file <- rstudioapi::getActiveDocumentContext()$path
output_file <- stringr::str_replace(current_file, '.Rmd', '.R')
knitr::purl(current_file, output = output_file)
file.edit(output_file)
```
### Load custom themes and functions
```{r, message=FALSE}
source('analysis/universal/themes.R') # themes for all project components

source('analysis/universal/functions.R') # themes for all project components
theme_set(theme_publication())

source('analysis/perturbseq/scripts/functions.R')  # functions for this component; may be empty
```

### Functions
```{r}
run_go <- function(gene_subset,
                   gene_table,
                   key = 'ENTREZID') {
  tic()
  
  bp <- enrichGO(gene_subset[[key]],
                 universe = gene_table[[key]],
                 keyType = key,
                 OrgDb = org.Hs.eg.db,
                 readable = TRUE,
                 ont = 'BP')
  
  mf <- enrichGO(gene_subset[[key]],
                 universe = gene_table[[key]],
                 keyType = key,
                 OrgDb = org.Hs.eg.db,
                 readable = TRUE,
                 ont = 'MF')
  
  cc <- enrichGO(gene_subset[[key]],
                 universe = gene_table[[key]],
                 keyType = key,
                 OrgDb = org.Hs.eg.db,
                 readable = TRUE,
                 ont = 'CC')
  toc()
  
  results <- bind_rows(list(bp@result %>% mutate(ontology = 'BP'),
                            mf@result %>% mutate(ontology = 'MF'),
                            cc@result %>% mutate(ontology = 'CC'))) %>% 
    as_tibble() %>% 
    mutate(log10q = -log10(qvalue)) %>% 
    dplyr::select(ontology, ID, Description, Count, GeneRatio, BgRatio, pvalue, p.adjust, qvalue, log10q, everything())
  
  list('results' = results,
       'bp' = bp,
       'mf' = mf,
       'cc' = cc)
  
}
```



### Directories
```{r, message=FALSE}
analysis_dir <- 'analysis/perturbseq/output/full_heatmap' # analysis file output directory
data_dir <- 'data/perturbseq/derived/seurat' # data file output directory
figures_dir <- 'figures/perturbseq'

dir.create(analysis_dir, showWarnings = FALSE)
dir.create(data_dir, showWarnings = FALSE)
```

### Load libraries
```{r, message=FALSE}
library(tidyverse)
library(Seurat)
library(pheatmap)
library(clusterProfiler)
library(org.Hs.eg.db)
library(sp)
```

## Import object
```{r}
integrated <- read_rds('data/perturbseq/derived/seurat/unified_qc.rds')
```

## Coordinates and metadata
```{r}
DIR <- 'analysis/perturbseq/output/data_export/'
pca_file <- 'pca_coordinates.tsv.gz'
umap_file <- 'umap_coordinates.tsv.gz'
meta_file <- 'coordinate_metadata.tsv.gz'

dr <- list('pca' = read_tsv(file.path(DIR, pca_file)),
           'umap' = read_tsv(file.path(DIR, umap_file)))
           
meta <- read_tsv(file.path(DIR, meta_file))
```


### Cell state hclust
```{r}
coordinates_hclust <- read_tsv('analysis/perturbseq/output/density_clustering/coordinates_hclust.tsv.gz')
coordinates_hclust_unique <- coordinates_hclust %>% dplyr::select(x, y, hc_cluster) %>% unique()

merged_clusters <- read_tsv('analysis/perturbseq/output/cell_state_umap/merged_clusters.tsv.gz')
```

```{r}
centers <- merged_clusters %>% dplyr::group_by(hc_cluster) %>% dplyr::summarize(x = median(UMAP_1), y = median(UMAP_2))


merged_clusters %>% 
  ggplot(aes(x = UMAP_1,
             y = UMAP_2,
             color = as.factor(hc_cluster))) +
  geom_point(alpha = 0.1,
              size = 0.1) +
  scale_color_manual(values = 
                       c('#0F8554', #1
                         '#6F4070', #2
                         '#0F8554', #3
                         '#1D6996', #4
                         '#E17C05', #5
                         '#38A6A5', #6
                         '#5F4690', #7
                         '#EDAD08', #8
                         '#0F8554', #9
                         '#38A6A5', #10
                         '#E17C05', #11
                         '#1D6996', #12
                         '#94346E', #13
                         '#5F4690', #14
                         '#73AF48', #15
                         '#5F4690', #16
                         '#38A6A5', #17
                         '#5F4690', #18
                         '#5F4690', #19
                         '#E17C05', #20
                         '#0F8554', #21
                         '#1D6996', #22
                         '#38A6A5', #23
                         '#EDAD08', #24
                         '#0F8554', #25
                         '#0F8554', #26
                         '#1D6996', #27
                         '#38A6A5', #28
                         '#EDAD08' #29
                         )) +
  shadowtext::geom_shadowtext(data = centers, aes(x = x, y = y, label = hc_cluster), color = 'white', size = 5, bg.color = 'black') +
  theme_void() +
  theme(legend.position = 'none')
```
```{r}
coordinates_hclust_unique
```
```{r}
all_hc_clusters <- coordinates_hclust %>% pull(hc_cluster) %>% unique()
all_hc_clusters
```

```{r eval=FALSE}

tic()
cell_state_barcodes <-  merged_clusters %>% filter(hc_cluster %in% all_hc_clusters)

cell_state_markers <- mclapply(all_hc_clusters, 
                                 mc.cores = 28,
                                 function(i) {
  
  selected_barcodes <- merged_clusters %>% mutate(selected = ifelse(hc_cluster == i, 1, 0))
  
  # find markers
  integrated_temp <- AddMetaData(integrated, selected_barcodes %>% column_to_rownames('barcode'))
  integrated_temp <- SetIdent(integrated_temp, value = 'selected')
  
  markers_temp <- FindMarkers(integrated_temp, ident.1 = 1, ident.2 = 0) %>%
    rownames_to_column('feature_id') %>% 
    mutate(hc_cluster = i)
  
  markers_temp
  
}) %>% 
  bind_rows() %>% 
  as_tibble() %>% 
  mutate(pct_diff = pct.1 - pct.2) %>% 
  dplyr::select(feature_id, hc_cluster, everything())

toc()

cell_state_markers %>% write_tsv(file.path(analysis_dir, 'full_cell_state_markers.tsv.gz'))



```

```{r}
cell_state_markers <- read_tsv(file.path(analysis_dir, 'full_cell_state_markers.tsv.gz'))
cell_state_markers
```


```{r}
cluster_subset <- c(16, 12, 13, 6, 18, 25, 20, 27, 1) %>% as.character()
markers_selected <- cell_state_markers %>% 
  filter(hc_cluster %in% cluster_subset) %>% 
  filter(str_detect(feature_id, '^MT', negate = T),
         str_detect(feature_id, '^RP', negate = T),
         str_detect(feature_id, '^MRP', negate = T),
         str_detect(feature_id, '^LH', negate = T)) %>% 
  group_by(feature_id) %>% 
  top_n(2, wt = avg_log2FC * pct_diff) %>% 
  group_by(hc_cluster) %>% 
  top_n(100, wt = abs(avg_log2FC) * pct_diff)
markers_selected
```
```{r}
selected_barcodes <- merged_clusters %>% filter(hc_cluster %in% c(18, 24)) %>% 
  mutate(selected = ifelse(hc_cluster == 18, 1, 0))
  
  # find markers
integrated_temp <- AddMetaData(integrated, selected_barcodes %>% column_to_rownames('barcode'))
integrated_temp <- SetIdent(integrated_temp, value = 'selected')

markers_temp <- FindMarkers(integrated_temp, ident.1 = 1, ident.2 = 0) %>%
  rownames_to_column('feature_id') %>% 
  as_tibble()

markers_temp
```

```{r}
cell_state_markers %>% filter(hc_cluster %in% c(18, 24)) %>% add_count(feature_id) %>% filter(n == 2)
```


### Pheatmap
```{r fig.height = 6, fig.width=6}
barcodes <- cell_state_barcodes %>% filter(hc_cluster %in% cluster_subset) %>% pull(barcode) %>% unique()
genes <-  c(markers_selected$feature_id, 'BAX') %>% unique()
data_mat <- integrated@assays$RNA@data[genes, barcodes]

data_table <- cell_state_barcodes %>%
  filter(hc_cluster %in% cluster_subset) %>% 
  inner_join(data_mat %>% as.matrix %>% t() %>% as.data.frame() %>% rownames_to_column('barcode')) %>% 
  dplyr::select(cluster = hc_cluster, all_of(genes)) %>% 
  group_by(cluster) %>% 
  summarize_all(mean) %>% 
  ungroup() %>% 
  mutate(group = paste0(cluster)) %>% 
  arrange(group) %>% 
  dplyr::select(-cluster) %>% 
  column_to_rownames('group')

pheat1 <- pheatmap(mat = t(data_table)[, cluster_subset],
                  scale = 'row',
                  clustering_distance_cols = 'correlation',
                  clustering_distance_rows = 'correlation',
                  clustering_method = 'average',
                  #clustering_method = 'complete',
                  cluster_cols = FALSE,
                  cluster_rows = TRUE,
                  filename = 'figures/perturbseq/full_cell_state_heatmap_full.pdf',
                  height = 20,
                  width = 6,
                  breaks = seq(-2, 2, length.out = 101),
                  color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(100),
                  fontsize_col = 8,
                  fontsize_row = 3,
                  angle_col = 0,
                  show_rownames = T)

## smaller
pheat2 <- pheatmap(mat = t(data_table)[, cluster_subset],
                  scale = 'row',
                  clustering_distance_cols = 'correlation',
                  clustering_distance_rows = 'correlation',
                  clustering_method = 'average',
                  #clustering_method = 'complete',
                  cluster_cols = FALSE,
                  cluster_rows = TRUE,
                  filename = 'figures/perturbseq/full_cell_state_heatmap_small.pdf',
                  height = 6,
                  width = 6,
                  breaks = seq(-2, 2, length.out = 101),
                  color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(100),
                  fontsize_col = 8,
                  angle_col = 0,
                  show_rownames = F)
```


### GO analysis
```{r}
expressed <- Matrix::rowSums(integrated@assays$RNA@counts > 0) > 0.01*ncol(integrated) # genes detected in at least 1% of cells

universe <- names(which(expressed))

lookup <- bitr(universe, 
               fromType = 'SYMBOL',
               OrgDb = org.Hs.eg.db,
               toType = c('ENTREZID', 'ENSEMBL'))

lookup
```

```{r}
markers_go_input <- cell_state_markers %>% 
  filter(str_detect(feature_id, '^MT', negate = T),
         str_detect(feature_id, '^RP', negate = T),
         str_detect(feature_id, '^MRP', negate = T),
         str_detect(feature_id, '^LH', negate = T)) %>% 
  filter(avg_log2FC > 0, p_val_adj < 0.01)

markers_go_input
```

```{r}

output_dir <- file.path(analysis_dir, 'go_results')
dir.create(output_dir, showWarnings = F)

completed_files <- list.files(output_dir) 
remaining_files <- markers_go_input$hc_cluster %>% unique() %>% setdiff(completed_files %>% str_remove('.tsv'))

tic()
# mclapply(remaining_files, 
#                      mc.cores = 28,
map(remaining_files,
    function(i) {
      
      cat(paste0('Cluster: ', i, '\n'))                     
      which_genes <- markers_go_input %>% filter(hc_cluster == i) %>% pull(feature_id)
      gene_subset <- lookup %>% filter(SYMBOL %in% which_genes)
      markers_go <- run_go(gene_subset, lookup)
      
      markers_go$results %>% mutate(cluster = i) %>% write_tsv(file.path(output_dir, paste0(i, '.tsv')))
      
    }) %>% invisible()
toc()


```

```{r}
completed_files <- list.files(output_dir) 
go_merge <- map(completed_files, function(i) {
  read_tsv(file.path(output_dir, i))
}) %>% bind_rows() %>% 
  dplyr::select(cluster, everything())

go_merge
```


```{r}
go_filtered <- go_merge %>% 
  filter(qvalue < 0.05) %>% 
  add_count(ID) %>% 
  group_by(cluster, ontology) %>% 
  arrange(cluster, qvalue) %>% 
  dplyr::select(cluster, n, ID, Description, Count, GeneRatio, BgRatio, qvalue, geneID) 

go_filtered 
```

```{r}
tic()
collapsed_go <- map(go_filtered$ID %>% unique(), 
                         function(i) {
                           associated_clusters <- go_filtered %>% filter(ID == i) %>% pull(cluster) %>% unique() %>% sort() %>% paste(collapse = ', ')
                           
                           tibble('ID' = i,
                                  'associated_clusters' = associated_clusters)
                         }) %>% bind_rows()
toc()

collapsed_go
```
```{r}
go_final <- go_filtered %>% 
  left_join(collapsed_go) %>% 
  dplyr::select(cluster, ontology, ID, n_clusters = n, associated_clusters, everything())

go_final
```


```{r}
barcodes <- meta %>%  pull(barcode) 
genes <-  full_markers_top$feature_id %>% unique()
data_mat <- integrated@assays$RNA@data[genes, barcodes]

data_table <- coordinate_metadata %>%

  inner_join(data_mat %>% as.matrix %>% t() %>% as.data.frame %>% rownames_to_column('barcode')) %>% 
  dplyr::select(cluster, all_of(genes)) %>% 
  group_by(cluster) %>% 
  summarize_all(mean) %>% 
  ungroup() %>% 
  mutate(group = paste0(cluster)) %>% 
  arrange(group) %>% 
  dplyr::select(-cluster) %>% 
  column_to_rownames('group')

pheatmap(mat = t(data_table),
         scale = 'row',
         clustering_distance_cols = 'correlation',
         clustering_distance_rows = 'correlation',
         clustering_method = 'ward.D2',
         cluster_rows = TRUE,
         color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdBu")))(100),
         fontsize = 7)
```

```{r}
tic()
collapsed_clusters <- map(markers_go_input$feature_id %>% unique(), 
                               function(i) {
                                 expressed_clusters <- markers_go_input %>% filter(feature_id == i) %>% pull(hc_cluster) %>% unique() %>% sort() %>% paste(collapse = ', ')
                                 
                                 tibble('feature_id' = i,
                                        'de_clusters' = expressed_clusters)
}) %>% bind_rows()
toc()
```

### Target cell state heatmap

### Density
```{r}
heatmap_subset <- coordinates_hclust %>% 
  left_join(meta %>% dplyr::select(target = guide_target, library, diff_hit_direction, diff_full_status, locus_ambiguity) %>% unique()) %>% 
  filter(library == 'Coding' | locus_ambiguity == 'Non-ambiguous', diff_full_status != 'Neighbor hit', diff_hit_direction != 'Non-hit')

cell_state_target_density <-  heatmap_subset %>% 
  group_by(target, hc_cluster) %>% 
  dplyr::summarize(density = quantile(z, probs = 0.25))

cell_state_target_density_wide <- cell_state_target_density %>% 
  dplyr::select(target, hc_cluster, density) %>% 
  arrange(hc_cluster) %>% 
  pivot_wider(names_from = 'hc_cluster', values_from = 'density', values_fill = 0) %>% 
  arrange(target) %>% 
  column_to_rownames('target')

cell_state_target_density_wide
```


### Density-based heatmap
```{r}
row_anno <- heatmap_subset %>% dplyr::select(target, library, diff_hit_direction) %>% unique() %>% arrange(target) %>% column_to_rownames('target')

diff_hit_direction_colors <- c('orangered', 'dodgerblue4')
names(diff_hit_direction_colors) <- c('Negative Hit', 'Positive Hit')

library_colors <- c('wheat3', 'darkorchid4')
names(library_colors) <- c('Coding', 'lncRNA')

anno_colors <- list(diff_hit_direction = diff_hit_direction_colors,
                    library = library_colors)

output_dir <- file.path(analysis_dir, 'cell_state_target_heatmap')
dir.create(output_dir)

order <- c(16, 12, 9, 24, 23, 
           21, 14, 3, 13, 6, 
           20, 27, 15, 10, 4, 18, 25, 8, 5, 
           2, 11, 19, 1, 7, 17, 28, 26 ,22, 29)
which_method <- 'average'
which_distance <- 'correlation'

pheat <- pheatmap(mat = cell_state_target_density_wide[, order],
                  annotation_row = row_anno,
                  scale = 'row',
                  clustering_distance_rows = which_distance,
                  clustering_method = which_method,
                  cluster_cols = FALSE,
                  cluster_rows = TRUE,
                  filename = 'figures/perturbseq/cell_state_target_heatmap_red.pdf',
                  color = colorRampPalette(colors = c('white', '#CC503E'))(100),
                  height = 10,
                  width = 8,
                  breaks = seq(0, 5, length.out = 101),
                  fontsize_col = 8,
                  fontsize_row = 4,
                  show_rownames = T,
                  legend = TRUE,
                  border_color = NA,
                  annotation_names_row = F,
                  annotation_names_col = F,
                  annotation_colors = anno_colors,
                  angle_col = 0)

pheat <- pheatmap(mat = cell_state_target_density_wide[, order],
                  annotation_row = row_anno,
                  scale = 'row',
                  clustering_distance_rows = which_distance,
                  clustering_method = which_method,
                  cluster_cols = FALSE,
                  cluster_rows = TRUE,
                  filename = 'figures/perturbseq/cell_state_target_heatmap_blue.pdf',
                  color = colorRampPalette(colors = c('white', 'dodgerblue4'))(100),
                  height = 10,
                  width = 8,
                  breaks = seq(0, 5, length.out = 101),
                  fontsize_col = 8,
                  fontsize_row = 4,
                  show_rownames = T,
                  legend = TRUE,
                  border_color = NA,
                  annotation_names_row = F,
                  annotation_names_col = F,
                  annotation_colors = anno_colors,
                  angle_col = 0)
```




```{r}

diff_hit_direction_colors <- c('orangered', 'dodgerblue4')
names(diff_hit_direction_colors) <- c('Negative Hit', 'Positive Hit')

library_colors <- c('red', 'blue')
names(library_colors) <- c('Coding', 'lncRNA')

anno_colors <- list(diff_hit_direction = diff_hit_direction_colors,
                    library = library_colors)


## smaller
pheat <- pheatmap(mat = cell_state_target_density_wide [clust_res$order, as.character(order)],
                  annotation_row = row_anno,
                  scale = 'row',
                  clustering_distance_cols = 'correlation',
                  clustering_distance_rows = 'correlation',
                  clustering_method = 'ward.D2',
                  cluster_cols = FALSE,
                  cluster_rows = FALSE,
                  height = 10,
                  width = 8,
                  border_color = NA,
                  breaks = seq(0, 5, length.out = 101),
                  color = colorRampPalette(c('white', 'dodgerblue4'))(100),
                  fontsize_col = 8,
                  fontsize_row = 4,
                  show_rownames = T,
                  legend = FALSE,
                  annotation_names_row = F,
                  annotation_names_col = F,
                  annotation_colors = anno_colors,
                  angle_col = 0)
```
```{r}
# compressed version for main figure
pheat <- pheatmap(mat = cell_state_target_density_wide [, as.character(order)],
                  annotation_row = row_anno,
                  scale = 'row',
                  clustering_distance_cols = 'correlation',
                  clustering_distance_rows = 'correlation',
                  clustering_method = 'ward.D2',
                  #clustering_method = 'complete',
                  cluster_cols = FALSE,
                  cluster_rows = TRUE,
                  #filename = 'figures/perturbseq/cell_state_target_heatmap.pdf',
                  height = 4,
                  width = 4,
                  border_color = NA,
                  breaks = seq(0, 5, length.out = 101),
                  color = colorRampPalette(c('white', 'dodgerblue4'))(100),
                  fontsize_col = 8,
                  fontsize_row = 4,
                  show_rownames = T,
                  legend = FALSE,
                  annotation_names_row = F,
                  annotation_names_col = F,
                  annotation_colors = anno_colors,
                  angle_col = 0)

pheat
```
## Export
```{r}
cell_state_target_density_wide %>% rownames_to_column('target') %>% write_tsv(file.path(analysis_dir, 'cell_state_target_density.tsv.gz'))
go_final %>% write_tsv(file.path(analysis_dir, 'go_final_full.tsv.gz'))
markers_go_input %>% 
  left_join(collapsed_clusters) %>% 
  write_tsv(file.path(analysis_dir, 'cell_state_markers_full_filtered.tsv.gz'))
```



## Session info
```{r}
sessionInfo()
```

