---
title: "Download MPRA Neural Induction Dataset"
output: html_notebook
---

## Purpose
Import data from Yosef/Ahituv study: https://www.sciencedirect.com/science/article/pii/S1934590919304217
Obtain ENA metadata table from: https://www.ebi.ac.uk/ena/browser/view/PRJNA473806

## Setup 
Set working directory to project directory
```{r setup}
require(knitr)
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

### Load custom themes and functions
```{r, message=FALSE}
source('analysis/universal/themes.R') # themes for all project components

source('analysis/universal/functions.R') # themes for all project components
theme_set(theme_publication())
```

### Directories
```{r, message=FALSE}
analysis_dir <- 'analysis/mpra/output/00_download' # analysis file output directory
data_dir <- 'data/mpra' # data file output directory

dir.create(analysis_dir, showWarnings = FALSE)
dir.create(data_dir, showWarnings = FALSE)
```

### Load libraries
```{r, message=FALSE}
library(tidyverse)
```

## Download
Load ENA metadata tables (split by assay)
Make sure to select columns including:
         library_name, 
         library_layout, 
         library_strategy, 
         library_selection, 
         experiment_title,
         sample_alias,
         sample_title,
         fastq_bytes,
         fastq_md5,
         fastq_aspera,
         sample_accession,
         experiment_accession,
         run_accession
```{r}
srp <- bind_rows(
  read_tsv('analysis/mpra/input/filereport_read_run_PRJNA473881_tsv.txt') %>% mutate(fastq_bytes = as.character(fastq_bytes)),
  read_tsv('analysis/mpra/input/filereport_read_run_PRJNA473882_tsv.txt'),
  read_tsv('analysis/mpra/input/filereport_read_run_PRJNA473883_tsv.txt'))

srp

```

Subset file
```{r}
sample_subset <- srp %>% 
  select( library_name, 
         library_layout, 
         library_strategy, 
         library_selection, 
         experiment_title,
         sample_alias,
         sample_title,
         fastq_bytes,
         fastq_md5,
         fastq_aspera,
         sample_accession,
         experiment_accession,
         run_accession) %>%
  separate(fastq_bytes, into = c('fastq1_bytes', 'fastq2_bytes'), sep = ';') %>% 
  separate(fastq_md5, into = c('fastq1_md5', 'fastq2_md5'), sep = ';') %>% 
  separate(fastq_aspera, into = c('fastq1_aspera', 'fastq2_aspera'), sep = ';') %>% 
  mutate(fastq1_gb = as.numeric(fastq1_bytes)/(1024^3),
         fastq2_gb = as.numeric(fastq2_bytes)/(1024^3)) %>% 
  separate(experiment_title, into = c(NA, 'experiment_title'), sep = ';') %>% 
  select(-fastq1_bytes, -fastq2_bytes)


sample_subset 
```


Calculate total fastq size
```{r}
sample_subset %>% pull(fastq1_gb) %>% sum(na.rm = TRUE)
```
```{r}
sample_subset %>% pull(fastq2_gb) %>% sum(na.rm = TRUE)
```

Prepare a bash script using ena-fast-download.py (https://github.com/wwood/ena-fast-download) that uses aspera-connect
Previously already set up aspera
```{r}
dir.create(file.path(data_dir, 'external/fastq'))
paste('ena-fast-download.py', sample_subset$run_accession) %>% 
  write(file.path(data_dir, 'external/fastq', 'aspera.txt'))
```

In case ena-fast-download is failing, set up manual aspera code
```{r}
paste0('ascp -QT -l 300m -P33001 -i /home/dwu/.aspera/connect/etc/asperaweb_id_dsa.openssh era-fasp@', c(sample_subset$fastq1_aspera, sample_subset$fastq2_aspera) %>% na.omit(), ' .') %>% 
  write(file.path(data_dir, 'external/fastq', 'aspera_manual.txt'))
```

Run script
```{bash eval=FALSE}
BASE=data/mpra/external
cd $BASE
cd fastq

while read LIST; do
$LIST; 
done < aspera_manual.txt
```

Check files
```{r}
all_files <- c(sample_subset$fastq1_aspera, sample_subset$fastq2_aspera) %>% na.omit() %>% basename() %>% str_remove('.fastq.gz')
all_files
```

```{r}
download_dir <- file.path(data_dir, 'external/fastq')
downloaded_files <- list.files(download_dir, pattern = '*.gz') %>% str_remove('.fq.gz') %>% str_remove('.fastq.gz')
downloaded_files
```

```{r}
missing_files <- setdiff(all_files, downloaded_files)
missing_files
```

## Export table
```{r}
sample_subset %>% write_tsv(file.path(analysis_dir, 'sample_table.tsv'))
```


## Session info
```{r}
sessionInfo()
```

