---
title: "CRISPRi sublibrary selection"
author: "David Wu"
output: html_notebook
---

## Purpose
Select lncRNA sublibraries for screen (not technically feasible to use all sublibraries)

## Setup 
Set working directory to project directory
```{r setup}
require(knitr)
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

### Load libraries
```{r, message=FALSE}
library(tidyverse)
```

### Load universal functions and themes
```{r}
source('analysis/rnaseq/scripts/functions.R') # functions for this subcomponent

source('analysis/universal/themes.R') # themes for all project components

theme_set(theme_publication())
```

### Output directory
```{r}
output_dir <- 'analysis/rnaseq/output/05_sublibrary'
dir.create(output_dir, showWarnings = FALSE)
```

### Load expression clustering data
```{r}
expression_clusters <- read_tsv('analysis/rnaseq/output/04_expression_clusters/expression_clusters.tsv.gz')
expression_clusters
```

### Load library data
```{r}
sgrna_library <- read_tsv('analysis/screen/input/crincl_reference/CRISPRi_v2_human_lincRNA_unique_merged_librarytable.txt')
sgrna_library
```

## Analysis
### Merge data
Combine expression data with sublibrary information, using mean expression and cluster number  
Allow duplicates for loci in multiple sublibraries
Include only lncRNAs (all coding sublibraries will be used)

```{r}
merged <- inner_join(expression_clusters %>% 
                       group_by(feature_id, cluster_fraction, cluster) %>% 
                       summarize(tpm = mean(tpm)),
                     sgrna_library %>% 
                       select(feature_id = gene,
                              sublibrary) %>% 
                       unique()) %>% 
  ungroup()

merged
```

### Select sublibraries
Feasible to assay ~10,000 loci
Prefer targets that are detected to undetected
```{r}
sublibraries <- merged %>% select(sublibrary, feature_id) %>% unique()
sublibraries
```
Summarize raw tallies of expression fraction of loci per sublibrary
```{r}
sublibrary_summary <- merged %>% 
  group_by(sublibrary, cluster_fraction) %>% 
  tally() %>% 
  pivot_wider(values_from = n,
              names_from = cluster_fraction) %>% 
  mutate(sublibrary_size = polyA + Total + `NA`,
         total_detected = polyA + Total,
         fraction_detected = round(total_detected/sublibrary_size, 2)) %>% 
  arrange(-fraction_detected) 

sublibrary_summary$cumulative <- cumsum(sublibrary_summary$sublibrary_size)

sublibrary_summary
```

Since some features are duplicated across libraries, select top libraries with at least 2/3 of features detected (~11k loci) and check if unique loci > 10,000
```{r}
selected_sublibraries <- sublibrary_summary %>% filter(fraction_detected > 2/3) %>% pull(sublibrary)
selected_sublibraries
```

Only ~9,000 loci
```{r}
merged %>% filter(sublibrary %in% selected_sublibraries) %>% select(feature_id) %>% unique()
```

Include most divergent library to cover more unique loci (Liu, Horlbeck et al Science 2017 supplementary figure S2A)
```{r}
selected_sublibraries <- c('cancer_common', 
                           'k562_hela_mcf7',
                           'u87_hek',
                           'u87_unique',
                           'hff_ips',
                           'ips_unique',
                           'common',
                           'hek_unique',
                           'u87_unique')
```

Check final count; close to 11k loci
```{r}
merged %>% filter(sublibrary %in% selected_sublibraries) %>% select(feature_id) %>% unique()
```


## Session info
```{r}
sessionInfo()
```


